---
title: "IL freely moving analysis"
author: "Viktor Plattner"
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: true
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
spacing: single
fontsize: 12pt
always_allow_html: yes
header-includes:
   - \usepackage{animate}

params:
  selected_animal: "GIIFM5"


knit: (
  function (inputFile, encoding) {
    
    rmarkdown::render(  
      input = inputFile,
      encoding = encoding,
      output_file = file.path(
        "notebook_html", 
        paste0("IL freely_moving_analysis_", 
        rmarkdown::yaml_front_matter(inputFile)$params$selected_animal,
        "_",
        Sys.Date(),
         ".html")
        ),
      envir = globalenv()
      )
    }
  )
---


<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
      font-family: "Arial", Helvetica, sans-serif;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #4d4e4f;
}
h1 { /* Header 1 */
  font-size: 28px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h2 { /* Header 2 */
  font-size: 22px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


```{r setup, include=FALSE}
### cheks all the used libraries in the project. Removes duplicates
### list of libraries mentioned in the R scripts of the project
library(tidyverse)

liblist <- map(
  list.files(pattern = ".Rmd|.R"),
  ~ readLines(con = .x) %>%
    grep(pattern = "library\\(", x = ., value = T)
) %>%
  unlist() %>%
  substr(
    start = 9,
    stop = regexpr(text = ., pattern = "\\)") - 1 %>% # regexpr: in case of multiple matches it only takes the first one
      as.vector()
  ) %>%
  `[`(!duplicated(.))

### list of installed packages
pkgs <- installed.packages() %>%
  as_tibble() %>%
  pull(1)

### packages that are needed but not installed
c(liblist[which(!liblist %in% pkgs)])

### installs uninstalled packages
install.packages(c(liblist[which(!liblist %in% pkgs)]))

### loads all libraries
lapply(liblist %>% as.list(), require, character.only = TRUE)




### defining global constants, filters and plot scales

library(imputeTS)

scale <- 1.6 ### pixel/mm
fps <- 29.5 ### frame rate (calculated from stim lengths in no. frames and known stim length in seconds)
standard_stim_length <- 10 ### this is the most commonly used stimulus length (in seconds)
around_stim_length <- as.integer(fps * 5)

time_res <- 1 / fps ### time between frames
ang_threshold <- 90
mouse_speed_mps <- 3.3



datapoint_to_plot <- 5

global_facet_theme <- theme(
  strip.text = element_text(size = 15),
  axis.text = element_text(size = 12),
  axis.title = element_text(size = 15)
)

xlim_mm <- c(0, 1280 / scale)
ylim_mm <- c(0, 720 / scale)

selected_animal <- params$selected_animal
selected_intensities <- c("1 mW", "5 mW", "10 mW")
control_animals <- c("GIIFM34b", "GIIFM35b", "GIIFM36b", "GIIFM37b")

`%!in%` <- Negate(`%in%`)




### custom functions used in the analysis



### removing junk animals from the data

RawDataFilter <- function(df) {
  # browser()
  df %>%
    dplyr::filter(
      stringr::str_detect(animal, pattern = "b")
    ) %>%
    as_tibble() %>%
    return()
}
```


```{r warning=F, message=F, echo=FALSE}
stimsIL <- read.csv(file.path("data", "IL_stim_rotation", "rotation_stims_IL.csv")) %>%
  as_tibble() %>%
  ### creating unique session IDs
  mutate(
    intensity = paste0(intensity, " mW"),
    # s2 = lag(session, default = "1"),
    # s3 = ifelse(s2 == session, 0, 1),
    stim_length = end - start
  ) %>%
  group_by(animal) %>%
  # mutate(
  #   session = cumsum(s3),
  #   session = paste0("session_", session)
  # ) %>%
  # select(-stim_duration, -s2, -s3) %>%
  rename("stim_number" = stim.number) %>%
  gather(key = "timing", value = "frame", start, end)%>% 
  ### no data file: video1coords.dat.9fix.bak, temporarily removed
  dplyr::filter(
    session != "video1coords.dat.9fix.bak"
  )

stimsIL %>%
  dplyr::filter(animal == selected_animal) %>%
  head(100)
```


```{r message=F, echo = FALSE, warning = FALSE}
coords_rawIL <- read.csv(file.path("data", "IL_stim_rotation", "rotation_all_IL.csv")) %>%
  as_tibble() %>%
  ### creating unique session IDs
  mutate(
    animal = gsub(pattern = ".*rotation", replacement = "", animal) %>%
      str_remove_all(pattern = "\\\\") %>% str_remove(pattern = "fix") %>% str_remove(pattern = "control") %>%
      substr(start = 1, stop = nchar(.) - 8) # ,
    # s2 = lag(session, default = "1"),
    # s3 = ifelse(s2 == session, 0, 1),
  ) %>%
  # group_by(animal) %>%
  # mutate(
  #   session = cumsum(s3),
  #   session = paste0("session_", session)
  # ) %>%
  # select(-s2, -s3) %>%
  # ungroup() %>%
  full_join(stimsIL) %>% 
  # group_by(animal) %>%
  # mutate(
  #   s2 = lag(session, default = "1"),
  #   s3 = ifelse(s2 == session, 0, 1),
  #   session = cumsum(s3),
  #   session = paste0("session_", session)
  # ) %>%
  # select(-s2, -s3) %>%
  # ungroup() %>%


  dplyr::group_by(animal, session) %>%
  dplyr::mutate(
    stim_number = ifelse(timing == "end", NA, stim_number)
  ) %>% 
  mutate(grp = cumsum(!is.na(stim_number))) %>%
  dplyr::group_by(animal, session, grp) %>%
  fill(timing) %>% 
  mutate(
    stim_freq = replace(stim_freq, timing == "start", stim_freq[1]),
    stim_freq = replace(stim_freq, timing == "end", NA),
    stim_length = replace(stim_length, timing == "start", stim_length[1]),
    stim_length = replace(stim_length, timing == "end", NA),
    intensity = replace(intensity, timing == "start", intensity[1]),
    intensity = replace(intensity, timing == "end", NA),
    stim_number = replace(stim_number, timing == "start", stim_number[1]),
    stim_number = replace(stim_number, timing == "end", NA),
    stim_duration = replace(stim_duration, timing == "start", stim_duration[1]),
    stim_duration = replace(stim_duration, timing == "end", NA),
    stim_side = replace(stim_side, timing == "start", stim_side[1]),
    stim_side = replace(stim_side, timing == "end", NA),
    quality = replace(quality, timing == "start", quality[1]),
    quality = replace(quality, timing == "end", NA),
    stim = F,
    stim = replace(stim, timing == "start", T)
  ) %>%
  ungroup() %>% 
  group_by(animal, session) %>%
  dplyr::mutate(
    rec_time = (frame / fps) %>% round(digits = 2),
    stim_categ = case_when(
      stim == T ~ "stim",
      lead(stim, n = around_stim_length) == T ~ "before stim",
      lag(stim, n = around_stim_length) == T ~ "after stim",
      stim == F ~ "no stim"
    ),

    ### shift values to the beginning of the stimulus block to be able to fill missing values in the next step
    stim_block = ifelse(stim_categ == "no stim", F, T),
    stim_number = lead(stim_number, n = around_stim_length),
    intensity = lead(intensity, n = around_stim_length),
    stim_freq = lead(stim_freq, n = around_stim_length),
    stim_side = lead(stim_side, n = around_stim_length),
    stim_duration = lead(stim_duration, n = around_stim_length),
    quality = lead(quality, n = around_stim_length),
  ) %>%  
  ungroup() %>%
  group_by(animal, session, stim_block) %>%
  fill(stim_number) %>%
  fill(intensity) %>%
  fill(stim_freq) %>%
  fill(stim_side) %>%
  fill(stim_duration) %>% 
  fill(quality) %>%
  # length of the stimulus
  unite(col = "head", head1, head2, sep = "_") %>%
  unite(col = "center", center1, center2, sep = "_") %>%
  unite(col = "tail", tail1, tail2, sep = "_") %>%
  gather(
    key = "body_part",
    value = "coordinate",
    head, center, tail
  ) %>%
  tidyr::extract(col = coordinate, into = c("X", "Y"), regex = "([[:alnum:]]+)_([[:alnum:]]+)") %>%
  dplyr::mutate(
    X = replace(X, miss == 0, NA),
    Y = replace(Y, miss == 0, NA)
  ) %>%
  dplyr::mutate(
    stim_number = replace(stim_number, is.na(stim_number), "no stim")
  ) 
  
```

```{r}
coordsIL <- coords_rawIL %>%
  dplyr::filter(
    body_part != "tail"
  ) %>%
  ### this block calculates distances and speed across frames
  dplyr::group_by(animal, body_part, session) %>%
  dplyr::mutate(

    ### interpolating missing values
    X_interp = imputeTS::na_interpolation(X %>% as.integer()) %>% round(),
    Y_interp = imputeTS::na_interpolation(Y %>% as.integer()) %>% round(),
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),

    ### sometimes the tracking software looses the point and jumps back and forth between unrealistic numbers or starts to follow something else
    ### these points are replaced by 0 (the animal did not change position between those frames)
    ### outliers are defined as Q1 - 3*IQR, Q3 + 3*IQR

    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / scale,
    speed_instant = d_mm / time_res,
    mmps = speed_instant * (1 / time_res),
    mps = (speed_instant * (1 / time_res)) / 1000,
    sec_counter = cumsum(ifelse(rec_time %% 1 == 0, 1, 0))
  ) %>%
  group_by(animal, session, body_part, sec_counter) %>%
  dplyr::mutate(speed_sec = sum(d_mm, na.rm = T)) %>%
  ungroup() %>% 
  dplyr::mutate(
    condition_grp = paste0(animal, " ", stim_side)
    # condition_grp = dplyr::case_when(
    #   ### bilat
    #   animal == "GIIFM5" & stim_side %in% c("bilat IL", "bilat Pf") ~ paste0(animal, "_bilat"),
    #   animal == "GIIFM6" & stim_side %in% c("bilat Pf") ~ paste0(animal, "_bilat"),
    #   ### unilat left
    #   animal == "GIIFM5" & stim_side %in% c("left IL", "left Pf") ~ "left",
    #   animal == "GIIFM6" & stim_side %in% c("left Pf") ~ "left",
    #   ### unilat right
    #   animal == "GIIFM5" & stim_side %in% c("right IL", "right Pf") ~ "right",
    #   animal == "GIIFM6" & stim_side %in% c("right IL", "right Pf") ~ "right"
    #   )
  ### adding categ_length: length of each stimulus category  
  ) %>% 
  dplyr::group_by(animal, body_part, session, stim_number, stim_categ) %>% 
  dplyr::mutate(
    categ_length = length(stim_categ)
    )

coordsIL %>%
  dplyr::filter(animal == selected_animal) %>%
  dplyr::group_by(session, body_part) %>%
  slice(1:5)
```





```{r message=F, warning=F, echo=F, include=F}
lengthsIL <- coords_rawIL %>%
  dplyr::filter(
    body_part == "head",
    stim_categ != "no stim",
  ) %>%
  group_by(animal, session, stim_number, intensity, quality) %>%
  count(stim_categ) %>%
  spread(stim_categ, n) %>%
  select(animal, session, stim_number, intensity, quality, `before stim`, stim, `after stim`)


```

```{r}
coordsIL %>% 
  dplyr::filter(stim_categ != "no stim") %>% 
  dplyr::group_by(animal, session) %>% 
  dplyr::summarise(unique_int_session = unique(intensity), num_unique_int_session = unique(intensity) %>% length())
```
```{r}
coordsIL %>% 
  dplyr::filter(stim_categ != "no stim") %>% 
  dplyr::group_by(animal, session) %>% 
  dplyr::summarise(unique_duration_session = unique(stim_duration), num_unique_duration_session = unique(stim_duration) %>% length())
```





<br>
Creating text layer to add intensity values to plots
<br>

```{r}
text_layer <- coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    stim == T
  ) %>%
  group_by(session) %>%
  summarise(int_label = intensity %>% unique())


text_layer_side <- coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    stim == T
  ) %>%
  group_by(session) %>%
  summarise(side_label = stim_side %>% unique())
```

<br>
Creating plot colors
<br>

```{r}
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n <- 6
cols <- gg_color_hue(n)
```



# Analysis (selected animal: `r selected_animal`)


## Checking tracking quality

Plotting X and Y positions vs time separately to see jumps in the tracking data and big separation between head-ceter-tail coordinates (unrealistically big mouse). The tracking software occasionally lost the animal and started to follow different points. If there are many of those the session cannot be analysed.


```{r warning=FALSE, fig.width=15, fig.height=10, echo = FALSE}

plt_xpos <- coords_rawIL %>%
  dplyr::filter(animal == selected_animal, body_part != "tail") %>%
  ggplot(mapping = aes(x = rec_time, y = X %>% as.integer())) +
  geom_line(aes(col = body_part)) +
  theme(legend.position = "none") +
  ylab("X position") +
  facet_grid(session ~ animal) +
  ggtitle(paste0("X position: ", selected_animal))



plt_ypos <- coords_rawIL %>%
  dplyr::filter(animal == selected_animal, body_part != "tail") %>%
  ggplot(mapping = aes(x = rec_time, y = Y %>% as.integer())) +
  geom_line(aes(col = body_part)) +
  theme(legend.position = "none") +
  ylab("Y position") +
  facet_grid(session ~ animal) +
  ggtitle(paste0("Y position: ", selected_animal))

cowplot::plot_grid(plt_xpos, plt_ypos, ncol = 2)
```


<br>
Plotting head-center distances over time. Dashed lines represent the outlier threshold (+- 1.5*IQR)
<br>


```{r fig.width=10, fig.height=40, echo = FALSE, message=FALSE, warning=FALSE}
coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    body_part != "tail"
  ) %>%
  dplyr::group_by(session, frame) %>%
  dplyr::mutate(
    dx = diff(X_interp),
    dy = diff(Y_interp),
    # ldx = lead(dx),
    # ldy = lead(dy),
    hc_d = sqrt(dx^2 + dy^2)
  ) %>%
  ungroup() %>%
  dplyr::filter(body_part == "center") %>%
  ggplot(
    mapping = aes(x = rec_time, y = hc_d)
  ) +
  geom_hline(
    aes(yintercept = quantile(hc_d, na.rm = T)["25%"] - 1.5 * IQR(hc_d, na.rm = T)),
    col = "gray50",
    lty = "dashed"
  ) +
  geom_hline(
    aes(yintercept = quantile(hc_d, na.rm = T)["75%"] + 1.5 * IQR(hc_d, na.rm = T)),
    col = "gray50",
    lty = "dashed"
  ) +
  geom_line(
    aes(col = stim, group = 1)
  ) +
  facet_wrap(~session, ncol = 1, scales = "free_y") +
  coord_cartesian(
    ylim = c(20, 130)
  ) +
  geom_label_repel(
    data = . %>%
      dplyr::filter(stim_number != "no stim", stim_categ == "stim") %>%
      dplyr::group_by(session, stim_number) %>%
      slice(1),
    aes(
      x = rec_time,
      y = mean(hc_d, na.rm = T),
      label = stim_duration
    ),
    direction = "x",
    nudge_y = 80
  ) +
  geom_label_repel(
    data = . %>%
      dplyr::filter(stim_number != "no stim", stim_categ == "stim") %>%
      dplyr::group_by(session, stim_number) %>%
      slice(1),
    aes(
      x = rec_time,
      y = mean(hc_d, na.rm = T),
      label = categ_length
    ),
    direction = "x",
    nudge_y = -30
  ) +
  global_facet_theme
```


```{r}
plt_taject_stim_baseline <- coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    # intensity == "5 mW",
    stim_categ != "no stim",
    stim == F,
    frame %% datapoint_to_plot == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / scale, 
      y = Y_interp, # / scale, 
      label = frame)
  ) +
  geom_path(
    aes(group = frame),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm"))
  ) +
  facet_grid(session~intensity) +
  xlim(c(0, 1280)) + # / scale)) +
  ylim(c(0, 720)) + # / scale)) +
  coord_fixed() + 
  # geom_label(
  #   data = text_layer %>% dplyr::filter(int_label %in% selected_intensities), 
  #   aes(
  #     x = Inf, y = -Inf, 
  #     hjust = 1.1, vjust = -0.5, 
  #     label = paste0("Intensity: ", int_label))) +
  labs(col = "Stimulus number") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(selected_animal, " baseline trajectory")) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  )


plt_taject_stim_stim <- coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    stim_categ != "no stim",
    #intensity == "5 mW",
    stim == T,
    frame %% datapoint_to_plot == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / scale, 
      y = Y_interp, # / scale, 
      label = frame)
  ) +
  
  geom_path(
    aes(group = frame, col = as.character(stim_number)),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm"))
  ) +
  facet_grid(session~intensity) +
  xlim(c(0, 1280)) + # / scale)) +
  ylim(c(0, 720)) + # / scale)) +
  coord_fixed() + 
  # geom_label(
  #   data = text_layer %>% dplyr::filter(int_label %in% selected_intensities), 
  #   aes(
  #     x = Inf, y = -Inf, 
  #     hjust = 1.1, vjust = -0.5, 
  #     label = paste0("Intensity: ", int_label))) +
  labs(col = "Stimulus number") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(selected_animal, " stimulus trajectory")) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  )


```


```{r fig.width=15, fig.height=30 }
cowplot::plot_grid(plt_taject_stim_baseline, plt_taject_stim_stim, nrow = 2)
```




```{r, message=F, warning=F,  fig.width=20,fig.height=10, message=FALSE, echo = FALSE}
coordsIL %>%
  dplyr::group_by(animal, body_part, session) %>%
  ggplot(mapping = aes(x = X_diff)) +
  stat_ecdf(
    aes(
      col = session %>% as.factor()
    ),
    geom = "point"
  ) +
  facet_grid(~animal ~ body_part) +
  global_facet_theme
```


```{r message=FALSE}
coordsIL %>%
  group_by(animal, body_part, session) %>%
  dplyr::filter(animal == selected_animal) %>%
  summarise(min = min(X_diff, na.rm = T), max = max(X_diff, na.rm = T))
```



```{r message=F, warning=F, fig.width=20, fig.height=10, echo = FALSE}
coordsIL %>%
  ggplot(mapping = aes(x = X_diff)) +
  geom_histogram(
    aes(
      fill = session %>% as.factor()
    ),
    binwidth = 5
  ) +
  scale_y_continuous(trans = "log10") +
  facet_grid(animal ~ body_part, scales = "free") +
  global_facet_theme
```



```{r fig.width=10, fig.height=10, echo = FALSE}
coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    session == "video0coords.dat.21fix.bak",
    body_part != "tail"
  ) %>%
  dplyr::group_by(session, frame) %>%
  # dplyr::arrange(body_part) %>%
  dplyr::mutate(
    dx = diff(X_interp),
    dy = diff(Y_interp),
    # ldx = lead(dx),
    # ldy = lead(dy),
    hc_d = sqrt(dx^2 + dy^2)
  ) %>%
  dplyr::group_by(body_part) %>%
  dplyr::slice(1:5) %>%
  ggplot(
    mapping = aes(x = X_interp, y = Y_interp, label = frame)
  ) +
  geom_line(
    aes(group = frame, col = frame %>% as.character()),
  ) +
  geom_point(
    aes(shape = body_part, col = frame %>% as.character()),
    size = 2
  ) +
  geom_label_repel(
    data = . %>% dplyr::filter(body_part == "head"),
    aes(
      x = X_interp + dx / 2,
      y = Y_interp + dy / 2,
      label = hc_d %>% round(4)
    ),
    direction = "x",
    nudge_x = -.7
  )
```



```{r}
line_cols <- gg_color_hue(3)

group_colors <- c("no stim" = "gray30", "before stim" = line_cols[1], "stim" = line_cols[3], "after stim" = line_cols[2])
group_colors2 <- c("no stim" = "gray30", "right before" = line_cols[1], "stim" = line_cols[3], "right after" = line_cols[2])
```



```{r message=F, fig.width=12, warning=F, echo = FALSE}

coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    body_part == "center",
    session == "video0coords.dat.21fix.bak",
    frame %% datapoint_to_plot == 0
  ) %>%
  ### recalculates distances after frames were removed
  dplyr::group_by(session) %>%
  dplyr::mutate(
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / scale,
    speed_instant = d_mm / time_res,
    mmps = speed_instant * (1 / time_res),
    mps = (speed_instant * (1 / time_res)) / 1000
  ) %>%
  dplyr::group_by(stim) %>%
  dplyr::mutate(mean_mps_roll = zoo::rollmean(x = d_mm, k = fps, fill = NA)) %>%
  ### plotting
  ggplot(aes(x = rec_time, y = d_mm)) +
  geom_point(aes(alpha = stim)) +
  geom_line(aes(x = rec_time, y = mean_mps_roll, col = stim_categ, group = 1), size = .8) +
  ggtitle(paste0(selected_animal, " mean traveled distance by stimulus period ", "(center)")) +
  scale_color_manual(values = group_colors)
```


```{r fig.width=12, fig.height=40, warning=FALSE, echo = FALSE}
coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    body_part == "center",
    frame %% datapoint_to_plot == 0
  ) %>%
  ### recalculates distances after frames were removed
  dplyr::group_by(session) %>%
  dplyr::mutate(
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / scale,
    speed_instant = d_mm / time_res,
    mmps = speed_instant * (1 / time_res),
    mps = (speed_instant * (1 / time_res)) / 1000
  ) %>%
  dplyr::group_by(session, stim_categ) %>%
  dplyr::mutate(mean_mps_roll = zoo::rollmean(x = d_mm, k = fps, fill = NA)) %>%
  ### plotting
  ggplot(aes(x = rec_time, y = d_mm)) +
  geom_point(aes(alpha = stim)) +
  geom_line(aes(x = rec_time, y = mean_mps_roll, col = stim_categ, group = 1), size = 1) +
  facet_wrap(~session, scales = "free", ncol = 1) +
  geom_label(data = text_layer, aes(x = -Inf, y = Inf, hjust = -.2, vjust = 1.3, label = paste0("Intensity: ", int_label))) +
  geom_label(data = text_layer_side, aes(x = Inf, y = Inf, hjust = 1.2, vjust = 1.3, label = paste0("Side: ", side_label))) +
    ggtitle(paste0(selected_animal, " mean traveled distance by stimulus period ", "(center)")) +
  scale_color_manual(values = group_colors) +
  global_facet_theme
```



```{r message=F, warning=F, echo=F}
coordsIL_1s_window <- bind_rows(
  ### before stim
  coordsIL %>%
    dplyr::filter(
      frame %% datapoint_to_plot == 0,
      stim_categ == "before stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / scale,
      speed_instant = d_mm / time_res,
      mmps = speed_instant * (1 / time_res),
      mps = (speed_instant * (1 / time_res)) / 1000
    ) %>%
    dplyr::group_by(animal, body_part, session, stim_categ, stim_number) %>%
    # dplyr::top_n(6),
    dplyr::slice(tail(row_number(), 6)),
  ### stim
  coordsIL %>%
    dplyr::filter(
      frame %% datapoint_to_plot == 0,
      stim_categ == "stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / scale,
      speed_instant = d_mm / time_res,
      mmps = speed_instant * (1 / time_res),
      mps = (speed_instant * (1 / time_res)) / 1000
    ) %>%
    dplyr::group_by(animal, body_part, session, stim_categ, stim_number) %>%
    dplyr::slice(1:6)
) %>% arrange(animal, body_part, session,rec_time)#, stim_number, stim_categ, rec_time)


coordsIL_1s_window %>%
  dplyr::filter(
    animal == selected_animal,
    intensity %in% selected_intensities
  ) %>%
  select(animal, session, frame, rec_time, stim_categ, d, stim_number, intensity, body_part, X_diff, Y_diff) %>%
  dplyr::group_by(animal, session, stim_categ) %>% 
  slice(1:6)
```





#### Calculating signed angles

```{r message=F, warning=F, echo = FALSE}

# angles <- coords %>%
#   dplyr::filter(
#     animal == selected_animal,
#     # frame %% datapoint_to_plot == 0, ### here it's good to have all the data points to calculate the cumulative angles
#     body_part != "tail"
#   ) %>%
#   dplyr::group_by(session, frame) %>%
#   mutate(
#     vectx = diff(X_interp),
#     vecty = diff(Y_interp)
#   ) %>%
#   dplyr::group_by(session) %>%
#   mutate(
#     leadx = lead(vectx),
#     leady = lead(vecty)
#   ) %>%
#   mutate(
#     signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
#   )

corrected_anglesIL <- coordsIL %>%
  dplyr::filter(
    animal == selected_animal,
    # frame %% datapoint_to_plot == 0, ### here it's good to have all the data points to calculate the cumulative angles
    body_part != "tail"
  ) %>%
  dplyr::group_by(session, frame) %>%
  mutate(
    vectx = diff(X_interp),
    vecty = diff(Y_interp)
  ) %>%
  dplyr::group_by(session) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>%
  mutate(
    signed_angle_corrected = case_when(
      signed_angle >= 180 ~ 360 - signed_angle,
      signed_angle %>% abs() < 180 ~ signed_angle,
      signed_angle <= -180 ~ -360 - signed_angle,
      signed_angle > -180 | signed_angle < 0 ~ signed_angle
    )
  ) %>%
  ### removing unrealistically large angles
  mutate(signed_angle_corrected = ifelse(signed_angle_corrected %>% abs() > ang_threshold, 0, signed_angle_corrected))
```



#### Plotting cumulative angles

Cumulative signed rotation angles during the sessions.

```{r message=FALSE,fig.width=10, fig.height=60}
corrected_anglesIL %>%
  group_by(session) %>%
  dplyr::filter(body_part == "center") %>%
  mutate(csum = cumsum(ifelse(is.na(signed_angle_corrected), 0, signed_angle_corrected)) %>% round(4)) %>%
  ggplot(
    mapping = aes(x = rec_time, y = csum %>% scale())
  ) +
  geom_line(aes(col = stim, group = 1), size = 1) +
  facet_wrap(~session, scales = "free_x", ncol = 1) +
  geom_label(data = text_layer, aes(x = -Inf, y = Inf, hjust = -.2, vjust = 1.3, label = paste0("Intensity: ", int_label))) +
  ggtitle(paste0(selected_animal, " cumulative angles by stimulus period ")) +
  global_facet_theme
```


# Population analysis

## Rotation

### Calculating signed population angles

```{r message=FALSE, warning=FALSE, echo = FALSE}
corrected_anglesIL_population <- coordsIL %>%
  dplyr::filter(
    body_part != "tail"
  ) %>%
  dplyr::group_by(animal, session, frame) %>%
  mutate(
    vectx = diff(X_interp),
    vecty = diff(Y_interp),
  ) %>%
  group_by(animal, session) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>%
  mutate(
    signed_angle_corrected = case_when(
      signed_angle >= 180 ~ 360 - signed_angle,
      signed_angle %>% abs() < 180 ~ signed_angle,
      signed_angle <= -180 ~ -360 - signed_angle,
      signed_angle > -180 | signed_angle < 0 ~ signed_angle
    )
  ) %>%
  ### removing unrealistically large angles
  mutate(
    signed_angle_corrected = ifelse(
      signed_angle_corrected %>% abs() > ang_threshold,
      0,
      signed_angle_corrected
    )
  )
  ### adding ipsi or contra rotation variable
  # dplyr::group_by(animal, intensity, stim_categ) %>%
  # dplyr::mutate(
  #   ipsicontra = case_when(
  #     sum(signed_angle_corrected, na.rm = T) > 0 & stim_side == "left" & stim_categ == "stim" ~ "ipsi",
  #     sum(signed_angle_corrected, na.rm = T) > 0 & stim_side == "right" & stim_categ == "stim" ~ "contra",
  #     sum(signed_angle_corrected, na.rm = T) < 0 & stim_side == "left" & stim_categ == "stim" ~ "contra",
  #     sum(signed_angle_corrected, na.rm = T) < 0 & stim_side == "right" & stim_categ == "stim" ~ "ipsi",
  #   )
  # )
```



```{r message=FALSE,fig.width=15, fig.height=60, out.width="100%"}
corrected_anglesIL_population %>%
  dplyr::filter(body_part == "head") %>%
  group_by(animal, session) %>%
  mutate(csum = cumsum(ifelse(is.na(signed_angle_corrected), 0, signed_angle_corrected)) %>% round(4)) %>%
  fill(intensity, .direction = "downup") %>%
  ggplot(
    mapping = aes(x = rec_time, y = csum %>% scale())
  ) +
  geom_line(aes(col = stim_categ, group = 1), size = 1) +
  # geom_text(
  #   data = coords$intensity %>% unique(),
  #   aes(label=intensity), check_overlap = T) +
  facet_grid(session~animal) + # factor(intensity, levels = gtools::mixedsort(intensity %>% unique())), scales = "free") +
  scale_color_manual(values = group_colors) +
  global_facet_theme +
  theme(
    axis.text.x = element_text(angle = 45),
  )
```





Mean/sum rotation angle by intensity (stimulus and control)

```{r message=FALSE, fig.height=20, fig.width=15, echo = FALSE}
sem <- function(x) sd(x) / sqrt(length(x)) # Create own function

plt1_angle <- corrected_anglesIL_population %>%
  dplyr::filter(
    animal %!in% c(control_animals, "GIIFM14b"),
    body_part == "head",
    stim_categ != "no stim",
    # stim_side == "left"
  ) %>%
  dplyr::filter(intensity %in% c(selected_intensities, "15 mW")) %>%
  dplyr::group_by(animal, intensity, stim_side, stim_categ) %>%
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected, na.rm = T),
    mean_ang = mean(signed_angle_corrected, na.rm = T),
    med_ang = median(signed_angle_corrected, na.rm = T),
    sd_ang = sd(signed_angle_corrected, na.rm = T),
    sem_ang = sem(signed_angle_corrected)
  ) %>%
  # dplyr::mutate(
  #   ipsicontra = case_when(
  #     sum_ang > 0 & stim_side == "left" & stim_categ == "stim" ~ "ipsi",
  #     sum_ang > 0 & stim_side == "right" & stim_categ == "stim" ~ "contra",
  #     sum_ang < 0 & stim_side == "left" & stim_categ == "stim" ~ "contra",
  #     sum_ang < 0 & stim_side == "right" & stim_categ == "stim" ~ "ipsi",
  #   )
  # ) %>%
  dplyr::group_by(animal, intensity) %>%
  # fill(ipsicontra, .direction = "up") %>%
# dplyr::filter(
#   sum_ang[stim_categ == "stim"] > 0) %>%
ggplot(
  mapping = aes(
    x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
    y = sum_ang %>% abs() /360  # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
  )
) +
  geom_point(aes(fill = animal), shape = 21, size = 4) +
  geom_line(aes(group = animal)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "less"),
    comparisons = list(c("before stim", "stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "greater"),
    comparisons = list(c("stim", "after stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  facet_grid(stim_side ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  global_facet_theme +
  expand_limits(y = c(0, 15)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  ggtitle("Stimulus") 
  #theme(aspect.ratio = 1) +
  # theme(
  #   axis.title.x = element_text(size = 12),
  #   axis.title.y = element_text(size = 12),
  #   axis.text = element_text(size = 12),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank(),
  #   panel.background = element_blank(),
  #   axis.line = element_line(colour = "black")
  # ) 


plt1_angle


``` 


```{r message=FALSE, fig.height=5, fig.width=15, echo = FALSE}
sem <- function(x) sd(x) / sqrt(length(x)) # Create own function

plt1_angle_categ_length_norm <- corrected_anglesIL_population %>%
  dplyr::filter(
    !is.na(condition_grp), 
    !str_detect(condition_grp, pattern = "bilat"),
    condition_grp %!in% c("GIIFM6 right IL left Pf", "GIIFM6 right IL right Pf"),
    body_part == "head",
    stim_categ != "no stim"
  ) %>% 
  dplyr::filter(intensity %in% c(selected_intensities, "15 mW")) %>%
  dplyr::group_by(condition_grp, intensity, stim_side, stim_categ) %>%
  
  ##### categ_length was created to animal groups not to condition groups!!!!!!!!!!!!!!!!!!! 
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected/categ_length, na.rm = T),
    mean_ang = mean(signed_angle_corrected/categ_length, na.rm = T),
    med_ang = median(signed_angle_corrected/categ_length, na.rm = T),
    sd_ang = sd(signed_angle_corrected/categ_length, na.rm = T),
    sem_ang = sem(signed_angle_corrected/categ_length)
  ) %>%
  dplyr::mutate(
    ipsicontra = case_when(
      sum_ang > 0 & str_detect(stim_side, pattern = "left") & stim_categ == "stim" ~ "ipsi",
      sum_ang > 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "left")  & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "ipsi"
    )
  ) %>% 
  dplyr::group_by(condition_grp, intensity) %>% 
  fill(ipsicontra, .direction = "up") %>% 
# dplyr::filter(
#   sum_ang[stim_categ == "stim"] > 0) %>%
ggplot(
  mapping = aes(
    x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
    y = sum_ang %>% abs()   # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
  )
) +
  geom_point(aes(fill = condition_grp), shape = 21, size = 4) +
  geom_line(aes(group = condition_grp)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "less"),
    comparisons = list(c("before stim", "stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "greater"),
    comparisons = list(c("stim", "after stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  facet_grid(ipsicontra ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  global_facet_theme +
  expand_limits(y = c(0, .05)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  ggtitle("Stimulus") +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )


plt1_angle_categ_length_norm


``` 

```{r message=FALSE, fig.height=5, fig.width=15, echo = FALSE}
sem <- function(x) sd(x) / sqrt(length(x)) # Create own function

plt1_angle <- corrected_anglesIL_population %>%
  dplyr::filter(
    !is.na(condition_grp), 
    !str_detect(condition_grp, pattern = "bilat"),
    condition_grp %!in% c("GIIFM6 right IL left Pf", "GIIFM6 right IL right Pf"),
    body_part == "head",
    stim_categ != "no stim"
  ) %>% 
  dplyr::filter(intensity %in% c(selected_intensities, "15 mW")) %>%
  dplyr::group_by(condition_grp, intensity, stim_side, stim_categ) %>%
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected, na.rm = T),
    mean_ang = mean(signed_angle_corrected, na.rm = T),
    med_ang = median(signed_angle_corrected, na.rm = T),
    sd_ang = sd(signed_angle_corrected, na.rm = T),
    sem_ang = sem(signed_angle_corrected)
  ) %>%
  dplyr::mutate(
    ipsicontra = case_when(
      sum_ang > 0 & str_detect(stim_side, pattern = "left") & stim_categ == "stim" ~ "ipsi",
      sum_ang > 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "left")  & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "ipsi"
    )
  ) %>% 
  dplyr::group_by(condition_grp, intensity) %>% 
  fill(ipsicontra, .direction = "up") %>% 
# dplyr::filter(
#   sum_ang[stim_categ == "stim"] > 0) %>%
ggplot(
  mapping = aes(
    x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
    y = (sum_ang/360) %>% abs()   # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
  )
) +
  geom_point(aes(fill = condition_grp), shape = 21, size = 4) +
  geom_line(aes(group = condition_grp)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "less"),
    comparisons = list(c("before stim", "stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "greater"),
    comparisons = list(c("stim", "after stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  facet_grid(ipsicontra ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  global_facet_theme +
  expand_limits(y = c(0, 15)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  ggtitle("Stimulus") +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )


plt1_angle


``` 


```{r fig.height=10, fig.width=15}
cowplot::plot_grid(plt1_angle,plt1_angle_categ_length_norm, nrow = 2)
```



```{r message=FALSE, fig.height=5, fig.width=15, echo = FALSE}
sem <- function(x) sd(x) / sqrt(length(x)) # Create own function

plt1_angle <- corrected_anglesIL_population %>%
  dplyr::filter(
    !is.na(condition_grp), 
    !str_detect(condition_grp, pattern = "bilat"),
    condition_grp %!in% c("GIIFM6 right IL left Pf", "GIIFM6 right IL right Pf"),
    body_part == "head",
    stim_categ != "no stim",
    stim_duration == 10
  ) %>%
  dplyr::filter(intensity %in% c(selected_intensities, "15 mW")) %>%
  dplyr::group_by(condition_grp, stim_side, stim_categ) %>% 
  
  
  #### categ_length was created to animal groups!!!!!
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected, na.rm = T),
    mean_ang = mean(signed_angle_corrected, na.rm = T),
    med_ang = median(signed_angle_corrected, na.rm = T),
    sd_ang = sd(signed_angle_corrected, na.rm = T),
    sem_ang = sem(signed_angle_corrected)
  ) %>% # dplyr::group_by(condition_grp, stim_number, stim_categ) %>% summarise(length(stim_number)) %>% View()
  dplyr::mutate(
    ipsicontra = case_when(
      sum_ang > 0 & str_detect(stim_side, pattern = "left") & stim_categ == "stim" ~ "ipsi",
      sum_ang > 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "left")  & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & str_detect(stim_side, pattern = "right") & stim_categ == "stim" ~ "ipsi"
    )
  ) %>% 
  dplyr::group_by(condition_grp) %>% 
  fill(ipsicontra, .direction = "up") %>%
  ungroup() %>% 
  dplyr::mutate(
    sum_ang_ipsiconra = ifelse(ipsicontra == "contra", sum_ang %>% abs(), -(sum_ang %>% abs())),
    mean_ang_ipsiconra = ifelse(ipsicontra == "contra", mean_ang %>% abs(), -(mean_ang %>% abs()))
  ) %>% # write.csv(file = file.path("output_data","to_publish","Emivel","IL-Pf_mean_No_rotation_per_second_all_intensities_ipsicontra.csv"))
# dplyr::filter(
#   sum_ang[stim_categ == "stim"] > 0) %>%
ggplot(
  mapping = aes(
    x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
    y = (mean_ang_ipsiconra * 29.5)/360  #sum_ang_ipsiconra/360 # sum_ang %>% abs() /360  # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
  )
) +
  geom_point(aes(fill = condition_grp), shape = 21, size = 4) +
  geom_line(aes(group = condition_grp)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "less"),
    comparisons = list(c("before stim", "stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "greater"),
    comparisons = list(c("stim", "after stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  #facet_grid(ipsicontra ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  global_facet_theme +
  geom_hline(yintercept = 0, col = "gray") +
  annotate(geom = "text", x = 3, hjust = -0.15, y = -0.02, label = "ipsilateral", col = "gray") +
  annotate(geom = "text", x = 3, hjust = -0.15, y = 0.02, label = "contralateral", col = "gray") +
  expand_limits(y = c(-0.2, 0.2), x = c(0,4)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  ggtitle("Stimulus") +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )


plt1_angle$data %>% 
  dplyr::group_by(stim_categ) %>% 
  summarise(
    mean((mean_ang_ipsiconra * 29.5)/360), 
    sd((mean_ang_ipsiconra * 29.5)/360), 
    median((mean_ang_ipsiconra * 29.5)/360))




``` 



## Traveled distance

```{r}
# saving the mean of the "before stim" "sum traveled distance" of all the animals (by intensity) for normalization

mean_sum_d_1mW <- coordsIL_1s_window %>%
  dplyr::filter(
    frame %% datapoint_to_plot == 0,
    body_part == "center",
    intensity %in% selected_intensities
  ) %>%
  dplyr::group_by(condition_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "1 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_5mW <- coordsIL_1s_window %>%
  dplyr::filter(
    frame %% datapoint_to_plot == 0,
    body_part == "center",
    intensity %in% selected_intensities
  ) %>%
  dplyr::group_by(condition_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "5 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_10mW <- coordsIL_1s_window %>%
  dplyr::filter(
    frame %% datapoint_to_plot == 0,
    body_part == "center",
    intensity %in% selected_intensities
  ) %>%
  dplyr::group_by(condition_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "10 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_15mW <- coordsIL_1s_window %>%
  dplyr::filter(
    frame %% datapoint_to_plot == 0,
    body_part == "center",
    intensity %in% c(selected_intensities, "15 mW")
  ) %>%
  dplyr::group_by(condition_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "15 mW"
  ) %>%
  pull(mean_sum_d)


coordsIL_1s_window %>%
  dplyr::filter(
    frame %% datapoint_to_plot == 0,
    body_part == "center",
    intensity %in% c(selected_intensities, "15 mW")
  ) %>%
  dplyr::group_by(condition_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(
    mean_sum_d = mean(sum_d),
    sd_sum_d = sd(sum_d)
  )
```





```{r fig.width=15, fig.height=5}
bind_rows(
  ### before stim 1 sec
  coordsIL_1s_window %>%
    dplyr::filter(
      stim_categ == "before stim",
      body_part == "center"
    ) %>%
    dplyr::mutate(
      stim_division = "before_stim",
      stim_div_length = 1
    ),
  ### stim devided to first_part (1s) and last part (9s)
  coordsIL %>%
    dplyr::filter(
      frame %% datapoint_to_plot == 0,
      stim_categ == "stim",
      body_part == "center"
      # animal %!in% control_animals,
      # body_part == "center",
      # intensity %in% selected_intensities
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / scale,
      speed_instant = d_mm / time_res,
      mmps = speed_instant * (1 / time_res),
      mps = (speed_instant * (1 / time_res)) / 1000
    ) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(animal, body_part, session, stim_side, stim_duration, stim_number) %>%
    dplyr::mutate(
      stim_division = ifelse(rec_time <= min(rec_time) + 1, yes = "first_part", no = "last_part"),
      # stim_div_length = ifelse(rec_time <= min(rec_time) + 1, yes = 1, no = 9)
    )
) %>%
  dplyr::group_by(animal, body_part, session, stim_side, stim_duration, stim_number, stim_division) %>%
  dplyr::mutate(stim_div_length = length(stim_division)) %>%
  ungroup() %>% 
  # arrange(animal, session, stim_number, rec_time) %>% View()



  dplyr::group_by(condition_grp, intensity, stim_categ , stim_division) %>% # stim_number, , stim_duration
  # dplyr::group_by(animal, body_part, intensity, stim_side) %>% 
  dplyr::summarise(
    sum_d = sum(d, na.rm = T),
    mean_d = mean(d, na.rm = T),
    max_sum_d = sum_d %>% max()
  ) %>% 
  dplyr::mutate(
    max_sum_d = replace(max_sum_d, stim_division != "before_stim", NA),
    mean_sum_d = case_when(
      intensity == "1 mW" ~ mean_sum_d_1mW,
      intensity == "5 mW" ~ mean_sum_d_5mW,
      intensity == "10 mW" ~ mean_sum_d_10mW,
      intensity == "15 mW" ~ mean_sum_d_15mW,
    )
  ) %>%
  dplyr::filter(
    intensity %in% c(selected_intensities, "15 mW")
  ) %>%
  ungroup() %>%
  fill(max_sum_d, .direction = "down") %>% 
  arrange(condition_grp, intensity) %>% 
  dplyr::filter(
    !is.na(condition_grp), 
    str_detect(condition_grp, pattern = "bilat| right IL left Pf")
  ) %>% 
  ggplot(
    mapping = aes(
      x = forcats::fct_relevel(stim_division, "before_stim", "first_part", "last_part"),
      y = mean_d # / stim_div_length #/mean_sum_d
    )
  ) +
  geom_point(aes(fill = condition_grp %>% as.factor()), shape = 21, size = 4) +
  facet_grid( ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  # facet_grid(animal ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  geom_line(aes(group = condition_grp)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    comparisons = list(c("before_stim", "first_part")),
    method.args = list(alternative = "less"),
    vjust = 1.6,
    hide.ns = F
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    comparisons = list(c("first_part", "last_part")),
    method.args = list(alternative = "greater"),
    vjust = 1.6,
    hide.ns = F
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    label.y = 55,
    paired = T,
    comparisons = list(c("before_stim", "last_part")),
    method.args = list(alternative = "less"),
    vjust = 1.7,
    hide.ns = F
  ) +
  expand_limits(y = c(0, 60)) +
  global_facet_theme +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  ) -> plt_pop_dist_stim #plt_pop_dist_ctrl

# sum_trav_dist_norm_by_before_mean

#cowplot::plot_grid(plt_pop_dist_stim, plt_pop_dist_ctrl, nrow = 2)

plt_pop_dist_stim
```








