---
title: "Freely moving analysis"
author: "Viktor Plattner"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: true
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: yes
spacing: single
fontsize: 12pt
always_allow_html: yes
---




<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
      font-family: "Arial", Helvetica, sans-serif;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #4d4e4f;
}
h1 { /* Header 1 */
  font-size: 28px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h2 { /* Header 2 */
  font-size: 22px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


# Loading data

## Defining global constants

scale: pixels/mm <br>
fps: frame rate <br>
time_res: time between frames <br>

```{r}
scale <- 1.38 ### pixel/mm
fps <- 30 ### frame rate
time_res <- 1 / fps ### time between frames
library(imputeTS)
```


## Loading stimulus times (frames)

```{r}
stims <- read.csv(file.path("data", "turning", "GIIFM19", "GIIFM19Frames_20201116.csv")) %>%
  as_tibble() %>%
  mutate(
    session = str_remove(session, "_"),
    stim_length = end - start
  ) %>%
  rename("stim_number" = stim.number) %>%
  gather(key = "timing", value = "frame", start, end)
stims
```

## Loading coordinates

```{r warning=FALSE, message=FALSE}
coords <- read.csv(file.path("data", "turning", "GIIFM19.csv")) %>%
  as_tibble() %>%
  mutate(session = str_remove(session, pattern = ".dat")) %>%
  full_join(stims) %>%
  dplyr::group_by(session) %>%
  mutate(grp = cumsum(!is.na(stim_number))) %>%
  dplyr::group_by(grp) %>%
  mutate(
    stim = F,
    stim = replace(stim, first(timing) == "start" | timing == "end", T)
  ) %>%
  # ungroup(grp) %>%
  group_by(stim) %>%
  fill(stim_number) %>%
  fill(intensity) %>%
  select(-grp, -animal, -timing, -stim_length) %>%
  #ungroup(stim) %>%
  group_by(session) %>%
  dplyr::mutate(
    rec_time = (frame / fps) %>% round(digits = 2),
    # segments = cumsum(ifelse(rec_time %% 9 == 0 & stim != T, 1, 0)),
    stim_categ = case_when(
      stim == T ~ "stim",
      lead(stim, n = 252) == T ~ "before stim",
      lag(stim, n = 252) == T ~ "after stim",
      stim == F ~ "no stim"
    )
  ) %>% # length of the stimulus
  unite(col = "head", head1, head2, sep = "_") %>%
  unite(col = "center", center1, center2, sep = "_") %>%
  unite(col = "tail", tail1, tail2, sep = "_") %>%
  gather(
    key = "body_part",
    value = "coordinate",
    head, center, tail
  ) %>%
  tidyr::extract(col = coordinate, into = c("X", "Y"), regex = "([[:alnum:]]+)_([[:alnum:]]+)") %>%
  dplyr::mutate(
    X = ifelse(
      as.integer(X) == 0 & as.integer(Y) == 0,
      yes = NA,
      no = as.integer(X)
    ),
    Y = ifelse(
      as.integer(X) == 0 & as.integer(Y) == 0,
      yes = NA,
      no = as.integer(Y)
    )
  ) %>%
  dplyr::group_by(session, body_part) %>%
  dplyr::mutate(
    # X = zoo::na.approx(X),
    # Y = zoo::na.approx(Y)
    X_interp = imputeTS::na.interpolation(X) %>% round(),
    Y_interp = imputeTS::na.interpolation(Y) %>% round(),
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / scale,
    speed_instant = d_mm / time_res,
    mmps = speed_instant * (1 / time_res),
    mps = (speed_instant * (1 / time_res)) / 1000,
    sec_counter = cumsum(ifelse(rec_time %% 1 == 0, 1, 0))
  ) %>%
  # mutate(
  #   X_diff = c(diff(X_interp), NA),
  #   Y_diff = c(diff(Y_interp), NA)
  # ) %>%
  # dplyr::mutate(
  #   d = sqrt(X_diff^2 + Y_diff^2),
  #   d_mm = d / scale
  # ) %>%
  # dplyr::mutate(
  #   speed_instant = d_mm / time_res,
  #   mmps = speed_instant * (1 / time_res),
  #   mps = (speed_instant * (1 / time_res)) / 1000,
  #   sec_counter = cumsum(ifelse(rec_time %% 1 == 0, 1, 0))
  # ) %>%
  group_by(session, sec_counter, body_part) %>%
  dplyr::mutate(speed_sec = sum(d_mm, na.rm = T)) %>%
  ungroup() %>%
  mutate(stim_number = replace(stim_number, is.na(stim_number), "no stim"))

coords
```


creating plot colors

```{r}
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n <- 5
cols <- gg_color_hue(n)
```

## Plots
### Plotting trajectories

```{r warning=FALSE}
coords %>%
  ggplot(
    mapping = aes(x = X, y = Y, label = frame)
  ) +
  geom_point(
    aes(shape = body_part, col = stim_number),
    size = 2,
    alpha = .1
  ) +
  geom_line(
    aes(group = frame, col = stim_number %>% as.character()),
    alpha = .1
  ) +
  scale_color_manual(
    values = c(cols[1], cols[2], cols[3], cols[4], cols[5], "black"),
    name = "Stimulus"
  ) +
  scale_shape(
    name = "Body part",
    labels = c("Center", "Head", "Tail")
  ) +
  xlim(c(0, 1280)) +
  ylim(c(0, 720)) +
  facet_wrap(~session)
```

```{r}
coords %>%
  dplyr::filter(body_part == "center") %>%
  group_by(session, sec_counter) %>%
  mutate(d_sec = sum(d_mm, na.rm = T)) %>%
  ggplot(
    mapping = aes(x = X, y = Y, label = frame)
  ) +
  geom_point(aes(color = d_sec %>% scale(), alpha = stim)) +
  # geom_line(aes(group = frame, col = stim_number %>% as.character())) +
  # scale_color_manual(
  #   values = c(cols[1], cols[2], cols[3], cols[4], cols[5], "black"),
  #   name = "Stimulus"
  # ) +
  # scale_shape(
  #   name = "Body part",
  #   labels = c("Center", "Head", "Tail")
  # ) +
  xlim(c(0, 1280)) +
  ylim(c(0, 720)) +
  facet_wrap(~session)
```

```{r}
coords %>%
  dplyr::filter(body_part == "center") %>%
  group_by(session, sec_counter) %>%
  mutate(d_sec = sum(d_mm, na.rm = T)) %>%
  ggplot(
    mapping = aes(x = X_interp, y = Y_interp, label = frame)
  ) +
  geom_point(aes(color = miss %>% as.character(), shape = stim)) +
  xlim(c(0, 1280)) +
  ylim(c(0, 720)) +
  facet_wrap(~session) +
  scale_color_discrete(name = "Interpolated values")

```




### Distribution of the distance between points on sequential frames

```{r}
coords %>%
  ggplot(mapping = aes(x = X_diff)) +
  geom_histogram() +
  facet_wrap(~body_part) + 
  scale_y_continuous(trans='log10') 
```
```{r}
coords %>%
  ggplot(mapping = aes(x = d_mm)) +
  geom_histogram() +
  facet_wrap(~body_part) +
  scale_y_continuous(trans='log10') 
  
```


### Traveled distance

```{r}
coords %>%
  dplyr::filter(body_part == "center") %>%
  ggplot(mapping = aes(x = rec_time, y = d_mm)) +
  geom_line(aes(color = stim, group = 1)) +
  geom_line(aes(x = rec_time, y = Y_diff -130), alpha = .5) + 
  geom_line(aes(x = rec_time, y = X_diff + 130), alpha = .5) + 
  facet_wrap(~session, scales = "free_x")
```

```{r}
coords %>%
  dplyr::group_by(session, stim) %>%
  dplyr::mutate(mean_mps_roll = zoo::rollmean(x = d_mm, k = 30, fill = NA)) %>%
  ggplot(aes(x = rec_time, y = d_mm)) +
  ylim(c(0, 100)) +
  geom_point(aes(color = stim)) +
  geom_line(aes(x = rec_time, y = mean_mps_roll)) +
  facet_wrap(~session, scales = "free_x")
```

## Calculating angles

### Toy data

```{r}
vectors <- tibble(
  body = c("cent", "h", "cent", "h", "cent", "h", "cent", "h"),
  frame = c(0, 0, 1, 1, 2, 2, 3, 3),
  X = c(750, 760, 770, 770, 755, 765, 755, 765),
  Y = c(500, 500, 500, 520, 510, 520, 510, 500)
)
vectors %>% arrange(body)
```


```{r}
ang <- function(a, b) {
  theta <- acos(sum(a * b) / (sqrt(sum(a * a)) * sqrt(sum(b * b))))
  return(theta * 180 / pi)
}

#ang(a = c(10, 0), b = c(10, 10))

vectors %>%
  group_by(frame) %>%
  summarise(dx = diff(X), dy = diff(Y)) %>%
  mutate(ldx = lead(dx), ldy = lead(dy)) %>%
  rowwise() %>%
  mutate(angle = ang(a = c(dx, dy), b = c(ldx, ldy))) %>%
  ungroup() %>%
  mutate(
    signed_angle = (atan2(dx, dy) - atan2(ldx, ldy)) * 180 / (pi)
  )
```


```{r}
matlib::angle(
  x = vectors %>%
    group_by(frame) %>%
    summarise(dx = diff(X), dy = diff(Y)) %>%
    dplyr::filter(frame == "0") %>%
    select(-1) %>% as.matrix() %>% as.vector(),
  y = vectors %>%
    group_by(frame) %>%
    summarise(dx = diff(X), dy = diff(Y)) %>%
    dplyr::filter(frame == "2") %>%
    select(-1) %>% as.matrix() %>% as.vector()
)
```


The dot product of two vectors x and y can be defined as:

$x \cdot y = ||x|| * ||y|| \cos(\theta) $

```{r}
vectors %>%
  ggplot(mapping = aes(x = X, y = Y)) +
  geom_point(aes(shape = body)) +
  geom_line(aes(group = frame, color = frame %>% as.character())) +
  geom_text_repel(aes(label = paste0(X, ",", Y), col = as.character(frame))) +
  xlim(750, 775) +
  ylim(500, 525)
```

## Real data (coords tibble)

```{r}
coords %>%
  dplyr::group_by(body_part) %>%
  slice(1:3) %>%
  dplyr::filter(body_part != "tail") %>%
  ggplot(
    mapping = aes(x = X, y = Y)
  ) +
  geom_point(aes(shape = body_part), size = 2) +
  geom_line(aes(group = frame, color = as.character(frame))) +
  geom_text_repel(aes(label = paste0(X, ",", Y), col = as.character(frame))) +
  geom_segment(aes(
    x = 210, xend = 255,
    y = 345, yend = 345
  ), linetype = "dashed", col = "red") +
  geom_segment(aes(
    x = 255, xend = 255,
    y = 345, yend = 364
  ), linetype = "dashed", col = "red")
```
```{r message=FALSE}
coords %>%
  dplyr::filter(body_part != "tail") %>%
  dplyr::group_by(session, frame) %>%
  summarise(
    vectx = diff(X_interp),
    vecty = diff(Y_interp)
  ) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  #rowwise() %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>%
   mutate(
    signed_angle2 = case_when(
        signed_angle >= 180 ~ 360 - signed_angle,
        signed_angle %>% abs() < 180 ~ signed_angle,
        signed_angle <= -180 ~ -360 - signed_angle,
        signed_angle > -180 | signed_angle < 0 ~ signed_angle
      )) %>% 
  ggplot(aes(x = signed_angle2)) +
  geom_histogram()
```

```{r message=F}
coords %>%
  dplyr::filter(body_part != "tail") %>%
  dplyr::group_by(session, frame) %>%
  #arrange(body_part) %>% 
  summarise(
    vectx = diff(X_interp),
    vecty = diff(Y_interp)
  ) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  #rowwise() %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>% 
  mutate(
    signed_angle2 = case_when(
        signed_angle >= 180 ~ 360 - signed_angle,
        signed_angle %>% abs() < 180 ~ signed_angle,
        signed_angle <= -180 ~ -360 - signed_angle,
        signed_angle > -180 | signed_angle < 0 ~ signed_angle
      )) %>% 

  add_column(
    rec_time = coords %>% dplyr::filter(body_part == "head") %>% pull(rec_time),
    stim = coords %>% dplyr::filter(body_part == "head") %>% pull(stim)
  ) %>%
  ggplot(aes(x = rec_time, y = signed_angle2)) +
  geom_line(aes(col = stim, group = 1)) +
  facet_wrap(~session) 
  #xlim(0,1)

```
```{r}
coords %>%
  dplyr::filter(session == "video0coords", body_part != "tail") %>% 
  dplyr::group_by(body_part) %>%
  dplyr::slice(1:25) %>% 
  ungroup() %>% 
  dplyr::group_by(frame) %>%
  summarise(
    vectx = diff(X_interp),
    vecty = diff(Y_interp)
  ) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  #rowwise() %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) 
  ggplot(
    mapping = aes(x = X, y = Y, label = frame)
  ) +
  geom_point(
    aes(shape = body_part, col = stim_number),
    size = 2
  ) +
  geom_line(
    aes(group = frame, col = stim_number %>% as.character())
  ) +
  scale_color_manual(
    values = c(cols[1], cols[2], cols[3], cols[4], cols[5], "black"),
    name = "Stimulus"
  ) +
  scale_shape(
    name = "Body part",
    labels = c("Center", "Head", "Tail")
  ) +
  xlim(c(0, 1280)) +
  ylim(c(0, 720)) 
```


```{r message=FALSE}
coords %>%
  dplyr::filter(body_part != "tail") %>%
  dplyr::group_by(session, frame) %>%
  summarise(
    vectx = diff(X),
    vecty = diff(Y)
  ) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  #rowwise() %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi) 
  ) %>%
   mutate(
    signed_angle2 = case_when(
        signed_angle >= 180 ~ 360 - signed_angle,
        signed_angle %>% abs() < 180 ~ signed_angle,
        signed_angle <= -180 ~ -360 - signed_angle,
        signed_angle > -180 | signed_angle < 0 ~ signed_angle
      )) %>% 
  add_column(
    rec_time = coords %>% dplyr::filter(body_part == "head") %>% pull(rec_time),
    stim = coords %>% dplyr::filter(body_part == "head") %>% pull(stim),
    stim_categ = coords %>% dplyr::filter(body_part == "head") %>% pull(stim_categ),
    stim_number = coords %>% dplyr::filter(body_part == "head") %>% pull(stim_number),
    Y_diff = coords %>% dplyr::filter(body_part == "head") %>% pull(Y_diff),
    X_diff = coords %>% dplyr::filter(body_part == "head") %>% pull(X_diff),
    d_mm = coords %>% dplyr::filter(body_part == "head") %>% pull(d_mm),
  ) %>%
  group_by(session) %>%
  mutate(csum = cumsum(ifelse(is.na(signed_angle2), 0, signed_angle2)) %>% round(4)) %>%
 # dplyr::filter(session == "video2coords") %>%
  # group_by(session) %>%
  # gather(key = "stim_categ" , value = "stim_value", stim, before_stim) %>%
  ggplot(
    mapping = aes(x = rec_time, y = csum %>% scale())
  ) +
  geom_line(aes(col = stim, group = 1)) +
  # geom_line(aes(x = rec_time, y = d_mm %>% scale() -7), alpha = .5) + 
  # geom_line(aes(x = rec_time, y = Y_diff %>% scale() -7), alpha = .5) + 
  # geom_line(aes(x = rec_time, y = X_diff %>% scale() + 7), alpha = .5) +
  facet_wrap(~session, scales = "free_x")

# mapping = aes(x = rec_time, y = cumsum(ifelse(is.na(signed_angle), 0, signed_angle) %>% scale()  ))

# ungroup() %>%
# dplyr::group_by(session, stim, before_stim) %>%
# summarise(mean(signed_angle, na.rm = T))
# dplyr::group_by(stim) %>%
#   summarise(sum(signed_angle, na.rm = T))
```
