---
title: "Frontal cortex-driven glycinergic inhibition of the intralaminar thalamus"
author: "Viktor Plattner, Emilia Bosz, Laszlo Acsady"
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: true
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
spacing: single
fontsize: 12pt
always_allow_html: yes
---




<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
      font-family: "Arial", Helvetica, sans-serif;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #4d4e4f;
}
h1 { /* Header 1 */
  font-size: 28px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h2 { /* Header 2 */
  font-size: 22px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial", Helvetica, sans-serif;
  color: #2c3a3e;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>



```{r setup, include=FALSE}
library(R.matlab)
library(tidyverse)
library(reshape2)
library(ggrepel)
library(DT)
library(ggpubr)
library(styler)
library(formatR)

library(bspec) ### power spectrum
library(WaveletComp) ### wavelet
library(diptest) ### to test distribution uni/multimodality (ISI)
library(exactRankTests) ### wilcox.test for ties
library(WaveletComp) ### wavelet
library(signal)

library(STAR)
library(see) # for half violinplot
library(gghalves) # for half violinplot

source(file.path("supplementary_functions", "CreateRecTibble.R"))

knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 30))
```


<br><br>



# Firing rate of IL neurons during glycinergic fiber activation

In these experiments glycinergic fibers were photactivated while the firing rate of individual IL neurons was recorded. Some of the IL cells had very low baseline activity. In order to increase the firing rate of the IL neurons and to detect the effect of the glycinergic fiber activation a small tail pinch was applied. The firing rate of the recorded IL cells decreased during the glycinergic fiber activation. 


## Loading data


Loading AP and stimuli times from *stimulus* and from *baseline* data to **IL_stim_firing** and to **IL_baseline_firing** tables:

```{r, message=FALSE, warning=FALSE, tidy=TRUE}
IL_stim_firing <- CreateRecTibble(
  AP_times = read_csv(file.path("data", "IL_MFR", "stimulus", "AP_times.csv")),
  stim_times = read_csv(file.path("data", "IL_MFR", "stimulus", "stim_times.csv"))
)
IL_stim_firing %>% head()
```

```{r, message=FALSE, warning=FALSE}
IL_baseline_firing <- CreateRecTibble(
  AP_times = read_csv(file.path("data", "IL_MFR", "baseline", "AP_times.csv")),
  stim_times = read_csv(file.path("data", "IL_MFR", "baseline", "stim_times.csv"))
)
IL_stim_firing %>% head()
```

<br><br>

## Summary information 

Summary information of the *stimulus* and  *baseline* recording files:

* Variables:
  + file names (*file_name*)
  + number of channels in the raw recording files (*No_ch*)
  + number of APs (*No_AP_unit*, *No_AP_unit2*)
  + number of stimulus trains (*No_Stim*)
  + LFP sampling rate (*samp_rate_lfp*)
  + unit sampling rate (*samp_rate_unit*)
  + length of the recording (*rec_length*)
  + length of the stimulus trains (*No_trains*)
  + Are the lengths of the stimulus trains equal in the recording (*train_length_equal*)
  + starting time of the stimulus trains (*train_start*)
  + ending time of the stimulus trains (*train_end*)

```{r, message=FALSE,echo=FALSE }

info_stim <- read_csv(file.path("data", "IL_MFR", "stimulus", "file_info.csv"))
datatable(info_stim,
  caption = "Summary table of the stimulus recording files (info_stim)",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```

<br>

```{r,message=FALSE,echo=FALSE}
info_baseline <- read_csv(file.path("data", "IL_MFR", "baseline", "file_info.csv"))
datatable(info_baseline,
  caption = "Summary table of the baseline recording files (info_baseline)",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```
<br>

**CELL_INFO** table to store the cell categories. It was created manually using the information from the summary excel table (*Glicy_juxta-fm_exp_records_injection_sum.xls*)

* Variables
  + cell identification (*cell_id*). To find the recording file use the summary excel table
  + file names (*file_name*)
  + spontaneously active cells (*bl_activity*)
  + induced firing (*pinch*)
  + individually identified neurons (*ident*)
  + control cells (*control*)


```{r, echo=FALSE}
### CELL INFO to store categories (spontaneously active, induced firing, identified)
CELL_INFO <- info_baseline %>%
  select(file_name) %>%
  add_column(cell_id = substr(info_baseline$file_name, 1, 6), .before = "file_name") %>%
  dplyr::mutate(bl_activity = T) %>%
  dplyr::mutate(bl_activity = replace(
    bl_activity,
    .$cell_id == "cell05" |
      .$cell_id == "cell10" |
      .$cell_id == "cell12" |
      .$cell_id == "cell13" |
      .$cell_id == "cell14" |
      .$cell_id == "cell15" |
      .$cell_id == "cell21" |
      .$cell_id == "cell24",
    F
  )) %>%
  dplyr::mutate(pinch = T) %>%
  dplyr::mutate(pinch = replace(
    pinch,
    .$cell_id == "cell01" |
      .$cell_id == "cell06" |
      .$cell_id == "cell07" |
      .$cell_id == "cell08" |
      .$cell_id == "cell18" |
      .$cell_id == "cell19" |
      .$cell_id == "cell20" |
      .$cell_id == "cell21" |
      .$cell_id == "cell23" |
      .$cell_id == "cell24" |
      .$cell_id == "cell25" |
      .$cell_id == "cell26" |
      .$cell_id == "cell27" |
      .$cell_id == "cell28",
    F
  )) %>%
  dplyr::mutate(ident = T) %>%
  dplyr::mutate(ident = replace(
    ident,
    .$cell_id == "cell06" |
      .$cell_id == "cell07" |
      .$cell_id == "cell11" |
      .$cell_id == "cell12" |
      .$cell_id == "cell13" |
      .$cell_id == "cell14" |
      .$cell_id == "cell21" |
      .$cell_id == "cell22" |
      .$cell_id == "cell23" |
      .$cell_id == "cell24" |
      .$cell_id == "cell25" |
      .$cell_id == "cell26" |
      .$cell_id == "cell27",
    F
  )) %>%
  dplyr::mutate(control = F) %>%
  dplyr::mutate(control = replace(
    control,
    .$cell_id == "cell25" |
      .$cell_id == "cell26" |
      .$cell_id == "cell27" |
      .$cell_id == "cell28" |
      .$cell_id == "cell29",
    T
  )) %>%
  dplyr::mutate(position = "unknown") %>%
  dplyr::mutate(position = replace(
    position,
    .$cell_id == "cell01" |
      .$cell_id == "cell02" |
      .$cell_id == "cell03" |
      .$cell_id == "cell10" |
      .$cell_id == "cell11" |
      .$cell_id == "cell14" |
      .$cell_id == "cell15",
    "IL"
  )) %>%
  dplyr::mutate(position = replace(
    position,
    .$cell_id == "cell20" |
      .$cell_id == "cell21" |
      .$cell_id == "cell22",
    "transient"
  )) %>%
  dplyr::mutate(position = replace(
    position,
    .$cell_id == "cell08" |
      .$cell_id == "cell16" |
      .$cell_id == "cell17" |
      .$cell_id == "cell23" |
      .$cell_id == "cell24",
    "PF"
  ))

datatable(CELL_INFO,
  caption = "CELL_INFO table",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)

# IL_stim_firing$file_name %>% unique() %>% length()
# info_stim$file_name %>% length()
```


<br><br>

## Calculations
### Firing rates

```{r, include=FALSE}
### CALCULATING --------

IL_stim_firing <- left_join(IL_stim_firing,
  info_stim %>%
    select(
      file_name,
      rec_length,
      train_starts,
      train_ends,
      train_length
    ),
  by = "file_name"
)

### calculating stim start times
start_times <- IL_stim_firing %>%
  dplyr::group_by(file_name) %>%
  dplyr::summarise(train_starts = unique(train_starts), train_ends = unique(train_ends)) %>%
  dplyr::mutate(cell_id = substr(info_stim$file_name, 1, 6)) %>%
  dplyr::group_by(cell_id) %>%
  dplyr::summarise(train_starts = unique(train_starts)) %>%
  pull(train_starts) %>%
  strsplit(",") %>%
  set_names(substr(info_stim$file_name, 1, 6)) %>%
  lapply(as.numeric)

### calculating stim end times
end_times <- IL_stim_firing %>%
  dplyr::group_by(file_name) %>%
  dplyr::summarise(train_starts = unique(train_starts), train_ends = unique(train_ends)) %>%
  dplyr::mutate(cell_id = substr(info_stim$file_name, 1, 6)) %>%
  dplyr::group_by(cell_id) %>%
  dplyr::summarise(train_ends = unique(train_ends)) %>%
  pull(train_ends) %>%
  strsplit(",") %>%
  set_names(substr(info_stim$file_name, 1, 6)) %>%
  lapply(as.numeric)


cell_list <- substr(info_stim$file_name, 1, 6)
CELL_INFO$cell_id
```

```{r, include=FALSE}
### function to calculate the number of AP b/d/a stim. Arguments: data (IL_stim_firing table with rec_length and stimulus train information)
BDACalculator <- function(data, list) {
  cell_list <- list

  train_length <- data %>%
    dplyr::filter(substr(file_name, 1, 6) == cell_list) %>%
    select(train_length) %>%
    pull() %>%
    `[[`(1)



  if (
    (data %>%
      dplyr::filter(substr(file_name, 1, 6) == cell_list, signal_type == "AP") %>%
      select(unit_id) %>%
      unique() %>%
      pull() %>%
      length()) == 1
  ) {
    ### In case of one unit in the file:
    AP_times <- data %>%
      dplyr::filter(substr(file_name, 1, 6) == cell_list, signal_type == "AP") %>%
      select(signal_time) %>%
      pull()

    AP_numbers <- matrix(0, nrow = length(start_times[[cell_list]]), ncol = 3)
    colnames(AP_numbers) <- c("b", "d", "a")

    for (train_num in 1:length(start_times[[cell_list]])) {
      ### before
      AP_numbers[train_num, 1] <- length(AP_times[start_times[[cell_list]][train_num] - train_length
      < AP_times & AP_times < start_times[[cell_list]][train_num]])

      ### during
      AP_numbers[train_num, 2] <- length(AP_times[AP_times > start_times[[cell_list]][train_num] & AP_times < end_times[[cell_list]][train_num]])

      ### after
      AP_numbers[train_num, 3] <- length(AP_times[AP_times > end_times[[cell_list]][train_num] &
        AP_times < end_times[[cell_list]][train_num] + train_length])
    }

    melt(AP_numbers, varnames = c("train", "stim_cond"), value.name = "No_AP") %>%
      as.tibble() %>%
      add_column(cell_id = cell_list) %>%
      add_column(train_length = train_length)
  } else {
    ### in case of multiple units in the file:
    AP_times_1 <- data %>%
      dplyr::filter(substr(file_name, 1, 6) == cell_list, signal_type == "AP", unit_id == 1) %>%
      select(signal_time) %>%
      pull()

    AP_numbers_1 <- matrix(0, nrow = length(start_times[[cell_list]]), ncol = 3)
    colnames(AP_numbers_1) <- c("b", "d", "a")

    for (train_num in 1:length(start_times[[cell_list]])) {
      ### before
      AP_numbers_1[train_num, 1] <- length(AP_times_1[start_times[[cell_list]][train_num] - train_length
      < AP_times_1 & AP_times_1 < start_times[[cell_list]][train_num]])

      ### during
      AP_numbers_1[train_num, 2] <- length(AP_times_1[AP_times_1 > start_times[[cell_list]][train_num] &
        AP_times_1 < end_times[[cell_list]][train_num]])

      ### after
      AP_numbers_1[train_num, 3] <- length(AP_times_1[AP_times_1 > end_times[[cell_list]][train_num] &
        AP_times_1 < end_times[[cell_list]][train_num] + train_length])
    }


    AP_times_2 <- data %>%
      dplyr::filter(substr(file_name, 1, 6) == cell_list, signal_type == "AP", unit_id == 2) %>%
      select(signal_time) %>%
      pull()

    AP_numbers_2 <- matrix(0, nrow = length(start_times[[cell_list]]), ncol = 3)
    colnames(AP_numbers_2) <- c("b", "d", "a")

    for (train_num in 1:length(start_times[[cell_list]])) {
      ### before
      AP_numbers_2[train_num, 1] <- length(AP_times_2[start_times[[cell_list]][train_num] - train_length
      < AP_times_2 & AP_times_2 < start_times[[cell_list]][train_num]])

      ### during
      AP_numbers_2[train_num, 2] <- length(AP_times_2[AP_times_2 > start_times[[cell_list]][train_num] &
        AP_times_2 < end_times[[cell_list]][train_num]])

      ### after
      AP_numbers_2[train_num, 3] <- length(AP_times_2[AP_times_2 > end_times[[cell_list]][train_num] &
        AP_times_2 < end_times[[cell_list]][train_num] + train_length])
    }



    if (cell_list %>% substr(5, 6) %>% as.numeric() %>% nchar() == 1) {
      new_name <- paste0(
        "cell0",
        cell_list %>% substr(5, 6) %>% as.numeric() + 1
      )
    }

    if (cell_list %>% substr(5, 6) %>% as.numeric() %>% nchar() == 2) {
      new_name <- paste0(
        "cell",
        cell_list %>% substr(5, 6) %>% as.numeric() + 1
      )
    }

    bind_rows(
      melt(AP_numbers_1, varnames = c("train", "stim_cond"), value.name = "No_AP") %>%
        as.tibble() %>%
        add_column(cell_id = cell_list) %>%
        add_column(train_length = train_length),

      melt(AP_numbers_2, varnames = c("train", "stim_cond"), value.name = "No_AP") %>%
        as.tibble() %>%
        add_column(cell_id = new_name) %>%
        add_column(train_length = train_length)
    )
  } # else
} # foo
```

**b_d_a_MFR**: Calculating the number of APs -using a custom made function (*BDACalculator*)- before during and after the stimulus trains (*b_d_a_MFR*). 

```{r}
b_d_a_MFR <- lapply(CELL_INFO$cell_id, BDACalculator, data = IL_stim_firing) %>%
  bind_rows() %>%
  dplyr::mutate(FR = No_AP / train_length) %>%
  dplyr::group_by(stim_cond, cell_id) %>%
  dplyr::summarise(MFR = mean(FR))
```

<br>
**sd_mean_isi**: Calculating the baseline MFR of the recorded IL cells from the *IL_baseline_firing* table using a custom made function (*SDMeanISI*). The results are stored in the *sd_mean_isi* table.

```{r, include=FALSE}
IL_baseline_firing <- left_join(IL_baseline_firing,
  info_baseline %>%
    select(file_name, rec_length),
  by = "file_name"
) %>%
  dplyr::mutate(stim_cond = "baseline")

### function to calculate ISI mean and SD during baseline activity of the cells
SDMeanISI <- function(f_name) {


  # mean_isi <- IL_baseline_firing %>%
  #   dplyr::filter(file_name == f_name, signal_type == "AP") %>%
  #   select(signal_time) %>%
  #   pull() %>%
  #   diff() %>%
  #   mean() * 1000 %>%
  #   `names<-`(f_name)
  #
  # sd_isi <- IL_baseline_firing %>%
  #   dplyr::filter(file_name == f_name, signal_type == "AP") %>%
  #   select(signal_time) %>%
  #   pull() %>%
  #   diff() %>%
  #   sd() * 1000 %>%
  #   `names<-`(f_name)
  #
  # return(list(MFR = mean_isi,
  #             SDFR = sd_isi))

  tmp2 <- IL_baseline_firing %>%
    dplyr::filter(file_name == f_name, signal_type == "AP") %>%
    select(signal_time) %>%
    pull() %>%
    diff() %>%
    mean() * 1000 %>%
      tibble(mean_isi = .)
  tmp2 %>%
    dplyr::mutate(sd_isi = IL_baseline_firing %>%
      dplyr::filter(file_name == f_name, signal_type == "AP") %>%
      select(signal_time) %>%
      pull() %>%
      diff() %>%
      sd() * 1000) %>%
    dplyr::mutate(file_name = f_name)
}

sd_mean_isi <- lapply(info_baseline$file_name, SDMeanISI) %>%
  bind_rows() %>%
  as.tibble()

sd_mean_isi <- sd_mean_isi %>%
  dplyr::mutate(cell_id = substr(sd_mean_isi$file_name, 1, 6)) %>%
  dplyr::mutate(MFR = 1000 / mean_isi) %>%
  dplyr::mutate(SD_FR = 1000 / sd_isi) %>%
  dplyr::mutate(MFR_inc = MFR + SD_FR) %>%
  dplyr::mutate(MFR_dec = MFR - SD_FR) %>%
  # dplyr::mutate(MFR_inc = 1000/(mean_isi-sd_isi)) %>%
  # dplyr::mutate(MFR_dec = 1000/(mean_isi+sd_isi)) %>%
  dplyr::mutate(stim_cond = "baseline") %>%
  dplyr::mutate(MFR_dec = replace(MFR_dec, sd_mean_isi$MFR_dec < 0, 0))
```
<br>

### Ranks (strength of inhibition)
**cellranks**: Calculating ranks based on the activity change from "*baseline*" to "*during stimulus*". If the activity change is negative (decreased MFR) the asigned rank is negative, if it is positive (increased MFR) the assigned rank is positive.

<br>

Calculating the firing rate change during stimulus (photoactivation of the glycinergic fibers) compared to baseline:

<br><br><br>

$$\mathbf{activity\_change} = \frac{during\_MFR - base\_MFR}{base\_MFR} * 100$$
<br>

```{r}
cellranks <- b_d_a_MFR %>%
  dplyr::group_by(stim_cond) %>%
  dplyr::mutate(base_MFR = sd_mean_isi$MFR) %>%
  dplyr::mutate(activity_change = ((MFR - base_MFR) / base_MFR * 100) %>% round(2)) %>%
  dplyr::filter(stim_cond == "d") %>%
  dplyr::mutate(change_rank = ifelse(activity_change > 0,
    rank(activity_change),
    -rank(-activity_change)
  )) %>%
  ungroup() %>%
  left_join(CELL_INFO %>%
    select(cell_id, control, pinch, position),
  by = "cell_id"
  )
cellranks
```



**cellranks_before_stim**: Calculating ranks based on the activity change from "*before stimulus*" to "*during stimulus*". If the activity change is negative (decreased MFR) the asigned rank is negative, if it is positive (increased MFR) the assigned rank is positive.

<br>

Calculating the firing rate change during stimulus (photoactivation of the glycinergic fibers) compared to before stimulus:

<br><br><br>

$$\mathbf{activity\_change} = \frac{during\_MFR - before\_MFR}{before\_MFR} * 100$$
<br>



```{r}
cellranks_before_stim <- b_d_a_MFR %>%
  spread(key = stim_cond, value = MFR) %>%
  dplyr::mutate(activity_change = ((d - b) / b * 100) %>% round(2)) %>%
  dplyr::mutate(change_rank = ifelse(activity_change > 0,
    rank(activity_change),
    -rank(-activity_change)
  )) %>%
  left_join(CELL_INFO %>%
    select(cell_id, control, pinch, position),
  by = "cell_id"
  )
cellranks_before_stim
```



<br><br>

### Data to plot
**TO_PLOT**: Combining *b_d_a_MFR* (firing rate of 29 IL neurons b/d/a stim) with *sd_mean_isi* table (baseline firing rate of the same 29 neurons), joining with CELL_INFO containing important information of the cells (baseline activity, identified, pinched, control) and with *cellranks* containing the ranks asigned to each cells based on the changes in MFR during the stimulus compared to baseline.

```{r, message=FALSE, warning=FALSE}
TO_PLOT <- bind_rows(
  sd_mean_isi %>%
    select(MFR, cell_id, stim_cond),
  b_d_a_MFR
) %>%
  left_join(CELL_INFO %>%
    select(-file_name),
  by = "cell_id"
  ) %>%
  left_join(cellranks %>%
    select(cell_id, change_rank, activity_change),
  by = "cell_id"
  )

datatable(TO_PLOT,
  caption = "TO_PLOT table",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```
<br><br>

## Plotting

### Baseline vs. "before" stimulus activity

Comparison of baseline and "before" stimulus firing rates in the case of spontaneously active and sponteneously inactive (pinch) neurons. Spontaneously inactive neurons showed significantly higher MFR before stiulus compared to baseline.

```{r, fig.width=10, echo = FALSE, warning=FALSE, out.width = "100%"}
ggplot(
  data = TO_PLOT %>%
    dplyr::filter(
      control == F,
      stim_cond == "baseline" | stim_cond == "b"
    ),
  mapping = aes(
    x = forcats::fct_relevel(stim_cond, "baseline", "b"),
    y = MFR
  )
) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  geom_point(aes(fill = pinch),
    shape = 21,
    # fill = "#EB8104",
    # color = "white",
    size = 4
  ) +
  # stroke = 2) +
  geom_line(
    aes(group = cell_id, col = pinch)
  ) +
  theme(panel.background = element_rect(fill = 0)) +
  theme_minimal() +
  # theme(
  #   axis.text = element_text(size = 20),
  #   text = element_text(size = 20)
  # ) +
  xlab("Recording period") +
  facet_wrap(vars(pinch)) +
  ggpubr::stat_compare_means(method = "wilcox.test", paired = T) +
  geom_label_repel(
    data = TO_PLOT %>%
      dplyr::filter(
        control == F,
        stim_cond == "baseline"
      ),
    mapping = aes(label = cell_id),
    nudge_x = -.9,
    direction = "y",
    # hjust = 2,
    segment.size = .2
  )
```

<br><br>

### MFR before, during and after stimulus 

```{r, fig.width=10 ,echo=FALSE, warning=FALSE}
ggplot(
  data = TO_PLOT %>%
    dplyr::filter(
      control == F,
      stim_cond == "b" | stim_cond == "d" | stim_cond == "a"
    ),
  mapping = aes(
    x = forcats::fct_relevel(stim_cond, "b", "d", "a"),
    y = MFR
  )
) +
  theme(panel.background = element_rect(fill = 0)) +
  theme_minimal() +
  # theme(
  #   axis.text = element_text(size = 20),
  #   text = element_text(size = 20)
  # ) +
  ggtitle("Pinch", ) +
  xlab("Recording period") +
  ylim(0, 35) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  geom_point(
    shape = 21,
    fill = "#EB8104",
    # color = "white",
    size = 4
  ) +
  # stroke = 2) +
  geom_line(
    aes(group = cell_id),
    color = "#EB8104"
  ) +
  geom_label_repel(
    data = TO_PLOT %>%
      dplyr::filter(
        control == F,
        stim_cond == "a"
      ),
    mapping = aes(label = cell_id),
    nudge_x = 0.15,
    direction = "y",
    hjust = -0.5,
    segment.size = .2
  ) +
  facet_wrap(vars(pinch))
```
<br><br>

### Strength of inhibition

Firing rate change **during** stimulus (photoactivation of the glycinergic fibers) compared to **baseline**:

```{r, fig.width=10, echo=FALSE,warning=FALSE}
ggplot(
  data = cellranks,
  mapping = aes(x = change_rank, y = activity_change)
) +
  geom_point(data = cellranks %>%
    dplyr::filter(control == F), size = 3) +
  geom_point(data = cellranks %>%
    dplyr::filter(control == T), size = 3, col = "magenta") +
  ### label for negative values
  geom_text_repel(
    data = cellranks %>%
      dplyr::filter(activity_change < 0),
    aes(label = cell_id),
    direction = "y",
    angle = 90,
    hjust = .5,
    nudge_y = -2
  ) +

  ### label for positive values
  geom_text_repel(
    data = cellranks %>%
      dplyr::filter(activity_change > 0),
    aes(label = cell_id),
    direction = "x",
    angle = 0,
    hjust = .5,
    nudge_x = -1
  ) +
  geom_hline(yintercept = 0, col = "grey", lty = "dashed") +
  annotate("text", x = 27, y = -17, label = "Baseline activity", col = "gray") +
  geom_vline(xintercept = 0, col = "grey", lty = "dashed") +
  annotate("text", x = -2, y = 280, label = "Negative rank: FR decrease", col = "gray", angle = 90) +
  annotate("text", x = 2, y = 280, label = "Positive rank: FR increase", col = "gray", angle = 90) +
  theme_minimal() +
  # theme(
  #   axis.text = element_text(size = 20),
  #   text = element_text(size = 20)
  # ) +
  ylim(-max(cellranks$activity_change), max(cellranks$activity_change))
```

<br><br>

Firing rate change **during** stimulus (photoactivation of the glycinergic fibers) compared to **before** stimulus:

```{r, fig.width=10, echo=FALSE,warning=FALSE}

ggplot(
  data = cellranks_before_stim,
  mapping = aes(x = change_rank, y = activity_change)
) +
  geom_point(
    data = cellranks_before_stim %>%
      dplyr::filter(control == F), size = 6,
    aes(
      col = position,
      shape = pinch
    )
  ) +
  scale_color_manual(values = c("IL" = "green", "transient" = "blue", "PF" = "red", "unknown" = "black", "control" = "magenta")) +
  scale_shape_manual(values = c("TRUE" = "+", "FALSE" = "-")) +
  geom_point(data = cellranks_before_stim %>%
    dplyr::filter(control == T), size = 3, col = "magenta") +
  ### label for negative values
  geom_text_repel(
    data = cellranks_before_stim %>%
      dplyr::filter(activity_change < 0),
    aes(label = cell_id),
    direction = "y",
    angle = 90,
    hjust = .5,
    nudge_y = -2
  ) +

  ### label for positive values
  geom_text_repel(
    data = cellranks_before_stim %>%
      dplyr::filter(activity_change > 0),
    aes(label = cell_id),
    direction = "x",
    angle = 0,
    hjust = .5,
    nudge_x = -1
  ) +
  geom_hline(yintercept = 0, col = "grey", lty = "dashed") +
  annotate("text", x = 27, y = -17, label = "Baseline activity", col = "gray") +
  geom_vline(xintercept = 0, col = "grey", lty = "dashed") +
  annotate("text", x = -2, y = 280, label = "Negative rank: FR decrease", col = "gray", angle = 90) +
  annotate("text", x = 2, y = 280, label = "Positive rank: FR increase", col = "gray", angle = 90) +
  theme_minimal() +
  # theme(
  #   axis.text = element_text(size = 20),
  #   text = element_text(size = 20)
  # ) +
  xlab("Ranks") +
  ylab("Change in firing rate [%]") +
  ylim(-max(cellranks_before_stim$activity_change), max(cellranks_before_stim$activity_change))
```












<br><br>

Plotting the change in MFR from "*baseline*" to "*during stimulus*". Coloring based on the strength of the inhibition (*rank*)

```{r, echo=FALSE,warning=FALSE}
ggplot(
  data = TO_PLOT %>%
    dplyr::filter(
      control == F,
      stim_cond == "baseline" | stim_cond == "d"
    ),
  mapping = aes(
    x = forcats::fct_relevel(stim_cond, "baseline", "d", "a"),
    y = MFR
  )
) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  geom_point(aes(fill = change_rank),
    shape = 21,
    # fill = "#EB8104",
    # color = "white",
    size = 4
  ) +
  # stroke = 2) +
  geom_line(
    aes(group = cell_id, col = change_rank)
  ) +
  geom_label_repel(
    data = TO_PLOT %>%
      dplyr::filter(
        control == F,
        stim_cond == "baseline"
      ),
    mapping = aes(label = cell_id, col = change_rank),
    # nudge_x = -.1,
    direction = "y",
    hjust = 4,
    segment.size = .2
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  # theme(
  #   axis.text = element_text(size = 20),
  #   text = element_text(size = 20)
  # ) +
  xlab("Recording period") +
  labs(fill = "Rank", col = "Rank")
```






<br><br><br><br><br><br>


# Activity of PRF glycinergic cells during PFC photoactivation

In these experiments we tested the effect of cortical photo-activation on the firing of PRF glycinergic neurons.  


## Loading and transforming (hidden code) data


```{r loading data, message=FALSE, warning=FALSE}
source(file.path("supplementary_functions", "CreateRecTibble.R"))
RECORDINGS <- CreateRecTibble(
  AP_times = read_csv(file.path("data", "cortical_stim", "AP_times.csv")),
  stim_times = read_csv(file.path("data", "cortical_stim", "stim_times.csv"))
)


### CAREFUL WITH THIS!! It's generated by the "CreatePSTHTibble" function but the frequencies have to be changed manually, and it can create duplicates! I have to work on that!
STIM_RESULTS <- read_csv(file.path("output_data", "cortical_stim_analysis.csv"))
```


```{r transforming data, echo=FALSE, message=FALSE, warning=FALSE}
### adding stimulus number within train
RECORDINGS <- RECORDINGS %>%
  dplyr::mutate(stim_number = 0)

### calculating stimulus number within train
initial_value <- RECORDINGS$stim_freq[1] %>%
  `comment<-`("First value of stim_freq variable. When it changes stimulus counting restarts")

stim_counter <- 1 %>%
  `comment<-`("Counts stimuli in a train")
index <- 1 %>%
  `comment<-`("Tracks the position (index) of stim_freq")

RECORDINGS$stim_number[1] <- stim_counter
repeat{
  if (RECORDINGS$stim_freq[index + 1] == initial_value) {
    RECORDINGS$stim_number[index + 1] <- stim_counter + 1
    stim_counter <- stim_counter + 1
    index <- index + 1
  } else {
    initial_value <- RECORDINGS$stim_freq[index + 1]
    stim_counter <- 1
    RECORDINGS$stim_number[index + 1] <- stim_counter
    index <- index + 1
  }

  if (index == length(RECORDINGS$stim_number)) {
    break
  }
}

### replacing stim_number with NA at "AP"
RECORDINGS$stim_number[RECORDINGS$signal_type == "AP"] <- NA

### categorizing stimuli
RECORDINGS <- RECORDINGS %>%
  dplyr::mutate(stim_freq_categ = stim_freq) %>%
  dplyr::mutate(stim_freq_categ = replace(
    x = stim_freq_categ,
    list = (stim_freq_categ == 12 | stim_freq_categ == 8),
    values = 10
  )) %>%
  dplyr::mutate(stim_freq_categ = replace(
    x = stim_freq_categ,
    list = (stim_freq_categ == 18),
    values = 20
  ))
```

```{r calculating hist stats, echo=FALSE, message=FALSE, warning=FALSE}
source(file.path("supplementary_functions", "HistStats.R"))

STIM_RESULTS %>%
  dplyr::group_by(freq) %>%
  dplyr::summarise(n_times = length(unique(animal_id))) %>%
  ungroup() %>%
  as.matrix() %>%
  unname() -> rep_matrix

### set PSTH range
PSTH_range <- c(-0.01, 0.04)


hist_stats <- STIM_RESULTS %>%
  dplyr::filter(
    first_ap_reltimes > PSTH_range[1],
    first_ap_reltimes < PSTH_range[2]
  ) %>%
  dplyr::group_by(freq) %>%
  do(gp_hist = ggplot(
    data = .,
    mapping = aes(x = first_ap_reltimes)
  ) +
    xlim(PSTH_range[1], PSTH_range[2]) +
    geom_histogram(aes(fill = animal_id), binwidth = 0.001) +
    scale_fill_brewer(
      palette = "Set3",
      name = "Animal"
      # guide = FALSE
    ) +
    facet_wrap(~animal_id, ncol = 3) +
    geom_vline(xintercept = 0) +
    xlab("Time (s)") +
    ylab("Count") +
    ggtitle(paste0(.$freq, " Hz"))) %>%
  `[[`(2) %>%
  lapply(. %>% HistStats() %>% as.tibble()) %>%
  bind_rows() %>%
  add_column(
    freq = c(
      rep(rep_matrix[1, 1], rep_matrix[1, 2]),
      rep(rep_matrix[2, 1], rep_matrix[2, 2]),
      rep(rep_matrix[3, 1], rep_matrix[3, 2])
    ),
    .after = "animal_id"
  )
```


```{r calculating response prob, echo=FALSE, message=FALSE, warning=FALSE}
RESP_PROB <- left_join(
  STIM_RESULTS %>%
    dplyr::filter(
      first_ap_reltimes > PSTH_range[1],
      first_ap_reltimes < PSTH_range[2]
    ) %>%
    dplyr::group_by(animal_id, freq) %>%
    dplyr::summarise(No_APs = length(first_ap_reltimes)) %>%
    dplyr::rename(stim_freq = freq) %>%
    dplyr::mutate(stim_freq = replace(stim_freq, stim_freq == 18, 20)) %>%
    dplyr::mutate(stim_freq = replace(stim_freq, stim_freq == 8, 10)),

  RECORDINGS %>%
    dplyr::filter(signal_type == "stim", stim_freq_categ %in% c("1", "10", "20")) %>%
    dplyr::group_by(animal_id, stim_freq_categ) %>%
    dplyr::summarise(No_stim = length(signal_time)) %>%
    dplyr::mutate(stim_freq = as.numeric(stim_freq_categ)) %>%
    select(-stim_freq_categ)
) %>%
  add_column(No_APs_range = hist_stats %>%
    arrange(animal_id) %>%
    dplyr::mutate(freq = replace(freq, freq == 18, 20)) %>%
    dplyr::mutate(freq = replace(freq, freq == 8, 10)) %>%
    dplyr::rename(stim_freq = freq) %>% pull(range_count_sum)) %>%
  dplyr::mutate(resp_prob = (No_APs / No_stim) * 100) %>%
  dplyr::mutate(resp_prob_range = (No_APs_range / No_stim) * 100) %>%
  add_column(peak_latency = hist_stats %>%
    arrange(animal_id) %>%
    pull(peak)) %>%
  add_column(mean_latency = hist_stats %>%
    arrange(animal_id) %>%
    pull(mean))
```

```{r calculating AP count, echo=FALSE, message=FALSE, warning=FALSE}
APCount <- function(animal) {
  tmp_ap <- RECORDINGS %>%
    dplyr::filter(signal_type == "AP", animal_id == animal) %>%
    select(signal_time) %>%
    pull()

  stim_trains <- RECORDINGS %>%
    dplyr::filter(signal_type == "stim") %>%
    dplyr::mutate(
      lag_sfc = lag(stim_freq),
      tmp_var = ifelse(lag_sfc == stim_freq & !is.na(lag_sfc), 0, 1),
      cum_tmp_var = cumsum(tmp_var)
    ) %>%
    dplyr::group_by(animal_id, stim_freq, cum_tmp_var) %>%
    dplyr::mutate(train_number = group_indices()) %>%
    ungroup() %>%
    select(-lag_sfc, -tmp_var, -cum_tmp_var) %>%
    dplyr::filter(animal_id == animal) %>%
    dplyr::group_by(stim_freq_categ, train_number, stim_number) %>%
    dplyr::summarise(
      No_APs = length(tmp_ap[tmp_ap > signal_time & tmp_ap < signal_time + 0.05]),
      AP_times = paste(tmp_ap[tmp_ap > signal_time & tmp_ap < signal_time + 0.05], collapse = ","),
      range_lower = signal_time,
      range_upper = signal_time + 0.05
    ) %>%
    add_column(animal_id = animal, .before = 1) %>%
    ungroup()
  return(stim_trains)
}

APC <- lapply(RECORDINGS$animal_id %>% unique(), APCount) %>% bind_rows()
```


* List of tibbles used to store data:
  + RECORDINGS tibble: stores AP and stim time stamps, number of stimuli in each train, stimulus frequency categories (eg. 8 10 and 12 Hz belong to 10 Hz category)
  + STIM_RESULTS: stores the data for PSTHs (CAREFUL WITH THIS!! It's generated by the "CreatePSTHTibble" function but the frequencies have to be changed manually, and it can create duplicates! I have to work on that! If something doesn't look right on PSTHs or response probability plots, check this one first)
  + hist_stats: stores PSTH statistics calculated from ggplot 
  + RESP_PROB: response probability and latency data calculated from stim results and recordings
  + APC: number of APs after each stimulus in a 50 ms window

```{r echo=FALSE, message=FALSE, warning=FALSE, tidy=TRUE}
datatable(RECORDINGS,
  caption = "RECORDINGS tibble",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, tidy=TRUE}
datatable(STIM_RESULTS,
  caption = "STIM_RESULTS tibble",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, tidy=TRUE}
datatable(hist_stats,
  caption = "PSTH statistics (hist_stats tibble)",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
datatable(RESP_PROB,
  caption = "Response probabilities",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
datatable(APC,
  caption = "Number of APs folowing each stimulus",
  rownames = TRUE,
  options = list(pageLength = 50, scrollX = T, scrollY = "500px", dom = "t")
)
```


## Plotting

### Distribution of frist APs after cortical activation


```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}

source(file.path("supplementary_functions", "geom_flat_violin.R"))
# PSTH_range <- c(0, 0.05)
ggplot(
  data = STIM_RESULTS %>%
    dplyr::filter(
      first_ap_reltimes > PSTH_range[1],
      first_ap_reltimes < PSTH_range[2]
    ),
  mapping = aes(y = first_ap_reltimes, x = animal_id)
) +
  coord_flip() +
  gghalves::geom_half_violin(side = "r", trim = T, scale = "width") +
  # geom_dotplot(binaxis = "y",
  #              dotsize = 4,
  #              stackdir = "down",
  #              binwidth = 0.0002,method = "histodot",binpositions = "all",
  #              ) +
  # see::geom_violindot(binwidth = 0.0001,size_dots = 6,) +
  geom_jitter(position = position_jitter(0), shape = "|") +
  # geom_boxplot(width = .1, outlier.colour = NA, position = "dodge") +
  geom_hline(yintercept = 0) +
  facet_wrap(~freq)
```

### PSTH @ 1 Hz cortical activation

```{r echo=FALSE, fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
### main plot

### set PSTH range: see earlier (calculating hist stats chunk)

### select frequency to plot (1, 8, 18)
freq_to_plot <- 1 %>% as.character()

gp_hist <- ggplot(
  data = STIM_RESULTS %>%
    dplyr::filter(
      first_ap_reltimes > PSTH_range[1],
      first_ap_reltimes < PSTH_range[2],
      freq == freq_to_plot
      # animal_id == "GII_21"
    ),
  mapping = aes(x = first_ap_reltimes)
) +
  xlim(PSTH_range[1], PSTH_range[2]) +
  geom_histogram(
    # aes(fill = animal_id),
    binwidth = 0.001
  ) +
  scale_fill_brewer(
    palette = "Set3",
    name = "Animal"
    # guide = FALSE
  ) +
  facet_wrap(~animal_id, ncol = 3) +
  geom_vline(xintercept = 0) +
  xlab("Time (s)") +
  ylab("Count")



### lines
gp_hist +
  ### median
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = median,
      xend = median,
      y = y_max + 7,
      yend = y_max + 18
    ),
    size = 1
  ) +

  ### data range
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = quantile25 - 1.5 * IQR,
      xend = quantile25,
      y = y_max + 12.5,
      yend = y_max + 12.5
    )
  ) +

  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = quantile75, xend = quantile75 + 1.5 * IQR, y = y_max + 12.5, yend = y_max + 12.5)
  ) +

  ### IQR (box)
  geom_rect(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(xmin = quantile25, xmax = quantile75, ymin = y_max + 7, ymax = y_max + 18),
    color = "red",
    alpha = 0,
    inherit.aes = F
  ) +

  ### mean
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = mean, y = y_max + 4), size = 2
  ) +

  ### peak
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = peak, y = y_max + 4),
    shape = 25,
    size = 2,
    fill = "red"
  )
```

### PSTH @ 10 Hz cortical activation

```{r echo=FALSE, fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
### main plot

### set PSTH range: see earlier (calculating hist stats chunk)

### select frequency to plot (1, 8, 18)
freq_to_plot <- 8 %>% as.character()

gp_hist <- ggplot(
  data = STIM_RESULTS %>%
    dplyr::filter(
      first_ap_reltimes > PSTH_range[1],
      first_ap_reltimes < PSTH_range[2],
      freq == freq_to_plot
      # animal_id == "GII_21"
    ),
  mapping = aes(x = first_ap_reltimes)
) +
  xlim(PSTH_range[1], PSTH_range[2]) +
  geom_histogram(
    # aes(fill = animal_id),
    binwidth = 0.001
  ) +
  scale_fill_brewer(
    palette = "Set3",
    name = "Animal"
    # guide = FALSE
  ) +
  facet_wrap(~animal_id, ncol = 3) +
  geom_vline(xintercept = 0) +
  xlab("Time (s)") +
  ylab("Count")



### lines
gp_hist +
  ### median
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = median,
      xend = median,
      y = y_max + 7,
      yend = y_max + 18
    ),
    size = 1
  ) +

  ### data range
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = quantile25 - 1.5 * IQR,
      xend = quantile25,
      y = y_max + 12.5,
      yend = y_max + 12.5
    )
  ) +

  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = quantile75, xend = quantile75 + 1.5 * IQR, y = y_max + 12.5, yend = y_max + 12.5)
  ) +

  ### IQR (box)
  geom_rect(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(xmin = quantile25, xmax = quantile75, ymin = y_max + 7, ymax = y_max + 18),
    color = "red",
    alpha = 0,
    inherit.aes = F
  ) +

  ### mean
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = mean, y = y_max + 4), size = 2
  ) +

  ### peak
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = peak, y = y_max + 4),
    shape = 25,
    size = 2,
    fill = "red"
  )
```

### PSTH @ 20 Hz cortical activation

```{r echo=FALSE, fig.width=10, fig.height=8, message=FALSE, warning=FALSE}
### main plot

### set PSTH range: see earlier (calculating hist stats chunk)

### select frequency to plot (1, 8, 18)
freq_to_plot <- 18 %>% as.character()

gp_hist <- ggplot(
  data = STIM_RESULTS %>%
    dplyr::filter(
      first_ap_reltimes > PSTH_range[1],
      first_ap_reltimes < PSTH_range[2],
      freq == freq_to_plot
      # animal_id == "GII_21"
    ),
  mapping = aes(x = first_ap_reltimes)
) +
  xlim(PSTH_range[1], PSTH_range[2]) +
  geom_histogram(
    # aes(fill = animal_id),
    binwidth = 0.001
  ) +
  scale_fill_brewer(
    palette = "Set3",
    name = "Animal"
    # guide = FALSE
  ) +
  facet_wrap(~animal_id, ncol = 3) +
  geom_vline(xintercept = 0) +
  xlab("Time (s)") +
  ylab("Count")



### lines
gp_hist +
  ### median
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = median,
      xend = median,
      y = y_max + 7,
      yend = y_max + 18
    ),
    size = 1
  ) +

  ### data range
  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(
      x = quantile25 - 1.5 * IQR,
      xend = quantile25,
      y = y_max + 12.5,
      yend = y_max + 12.5
    )
  ) +

  geom_segment(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = quantile75, xend = quantile75 + 1.5 * IQR, y = y_max + 12.5, yend = y_max + 12.5)
  ) +

  ### IQR (box)
  geom_rect(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(xmin = quantile25, xmax = quantile75, ymin = y_max + 7, ymax = y_max + 18),
    color = "red",
    alpha = 0,
    inherit.aes = F
  ) +

  ### mean
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = mean, y = y_max + 4), size = 2
  ) +

  ### peak
  geom_point(
    data = hist_stats %>%
      dplyr::filter(freq == freq_to_plot) %>%
      dplyr::group_by(animal_id),
    mapping = aes(x = peak, y = y_max + 4),
    shape = 25,
    size = 2,
    fill = "red"
  )
```


### Response probability and latency



```{r echo=FALSE, fig.width=8, message=FALSE, warning=FALSE}
ggplot(
  data = RESP_PROB, # %>%
  # dplyr::filter(animal_id != "GII_20"),
  mapping = aes(
    x = as.factor(stim_freq),
    y = resp_prob_range
  )
) +
  geom_boxplot(alpha = 0, width = 0.2, lwd = 1, fatten = 1.2) +
  geom_point(
    aes(
      # fill = animal_id
    ),
    fill = "gray",
    color = "black",
    size = 3,
    shape = 21,
    # position = position_jitter(0.1)
  ) +
  # geom_line(aes(group = animal_id)) +
  xlab("Stimulus frequency") +
  ylab("Response probability") +
  geom_signif(
    comparisons = list(c("1", "10")),
    map_signif_level = TRUE,
    test = "mood.test",
    color = "gray"
  ) +
  geom_signif(
    comparisons = list(c("10", "20")),
    map_signif_level = TRUE,
    test = "mood.test",
    margin_top = .1,
    color = "gray"
  ) +
  geom_text_repel(
    aes(label = animal_id),
    direction = "y",
    hjust = -.5
  )
```


```{r echo=FALSE, fig.width=8, message=FALSE, warning=FALSE}
ggplot(
  data = RESP_PROB %>%
    gather(
      key = "latency_type",
      value = "latency_value",
      peak_latency,
      mean_latency
    ) %>%
    dplyr::filter(animal_id != "GII_20"),
  mapping = aes(
    x = as.factor(stim_freq),
    y = latency_value * 1000
  )
) +
  ### data visualization layers
  geom_point(
    aes(fill = latency_type),
    position = position_dodge(0.4),
    size = 3,
    color = "black",
    shape = 21
  ) +
  geom_boxplot(
    aes(color = latency_type),
    position = position_dodge(0.4),
    width = 0.3,
    alpha = 0,
    lwd = 1,
    fatten = 1.2
  ) +
  ylim(0, 25) +
  xlab("Stimulus frequency") +
  ylab("Peak latency") +
  scale_color_discrete(
    name = "Latency",
    breaks = c("mean_latency", "peak_latency"),
    labels = c("Mean", "Peak"),
    guide = guide_legend(override.aes = list(color = c(rep(NA, 2))))
  ) +
  scale_fill_discrete(
    name = "Latency",
    breaks = c("mean_latency", "peak_latency"),
    labels = c("Mean", "Peak")
  ) +
  theme(legend.position = "bottom")
```


### Number of APs after each stimulus in stimulus trains (50 ms window)


```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
ggplot(
  data = APC %>%
    dplyr::filter(stim_freq_categ %in% c("1", "10", "20"), stim_number < 11) %>%
    dplyr::group_by(animal_id, stim_freq_categ, stim_number) %>%
    dplyr::summarise(No_APs = sum(No_APs)) %>%
    ungroup() %>%
    dplyr::group_by(stim_freq_categ, stim_number) %>%
    dplyr::summarise(
      No_AP_mean = mean(No_APs),
      No_AP_sd = sd(No_APs),
      SEM = sd(No_APs) / sqrt(length(No_APs))
    ),
  mapping = aes(
    x = stim_number,
    y = No_AP_mean,
    colour = as.factor(stim_freq_categ),
    group = as.factor(stim_freq_categ)
  )
) +
  geom_point() +
  geom_line() +
  xlab("No. stimulus in train") +
  ylab("Mean no. APs accross animals") +
  labs(color = "Frequency") +
  ylim(0, 18) +
  scale_x_discrete(limits = as.character(c(1:10))) +
  theme(legend.position = "bottom")
```

```{r echo=FALSE, fig.width=10, fig.height=4, message=FALSE, warning=FALSE}
ggplot(
  data = APC %>%
    dplyr::filter(stim_freq_categ %in% c("1", "10", "20"), stim_number < 11) %>%
    dplyr::group_by(animal_id, stim_freq_categ, stim_number) %>%
    dplyr::summarise(No_APs = sum(No_APs)),
  mapping = aes(
    x = stim_number,
    y = No_APs,
    color = as.factor(animal_id),
    linetype = as.factor(animal_id)
  )
) +
  geom_point() +
  geom_line() +
  scale_x_discrete(limits = c(1:10), breaks = c(1:10)) +
  geom_label_repel(
    data = . %>%
      dplyr::filter(stim_number == 10),
    aes(label = animal_id),
    direction = "y",
    hjust = -2,
    alpha = .7
  ) +
  expand_limits(x = 12) +
  theme(legend.position = "none") +

  # xlim(1,10) +
  facet_wrap(~stim_freq_categ) +
  labs(linetype = "Animals", color = "Animals") +
  xlab("No. stimulus in train") +
  ylab("Number of APs")
```

<br><br><br><br><br><br>

# Spontaneous desynchronization of the FC slow oscillation 

## Loading data 

Loading mat files to *file_list* object

Example file: **08_right_PnO_t5_4083_baseline1_wake.mat"**

```{r,  echo=FALSE,warning=FALSE,message=FALSE}
file_list <- list.files(
  path = "data",
  pattern = "*.mat", full.names = F, recursive = F
)

```

Content of mat files:

* AP channel
  + **title:** channel name
  + **comment:** notes about the channel
  + **resolution:** 2.5e-06
  + **interval:** time between two data points in sec (5e-05)
  + **scale:** scaling factor (y scale), real = (data * scale) + offset
  + **offset:** scale offset
  + **units:** unit of the channel (volt)
  + **length:** number of APs
  + **items:** No. points easch AP is represented by
  + **trigger:**
  + **traces:**
  + **times:** times of the APs (first data point of the waveform)
  + **codes:** marker codes of the APs
  + **values:** matrix containing the waveforms of the APs (described by a given No. points)
 
* EEG channel
  + **title:** channel name
  + **comment:**
  + **interval:**
  + **scale:** scaling factor (y scale), real = (data * scale) + offset
  + **offset:** scale offset
  + **units:**
  + **start:**
  + **length:**
  + **values:**

```{r}
file_to_load <- file_list[[1]]
filename <- as.character(substring(file_to_load, 1, nchar(file_to_load) - 4))
raw.rec <- readMat(file.path("data", file_to_load))

### takes the first AP (first row) and tells the index of point with the max value
points_to_peak <- which(raw.rec$ap[, , 1]$values[1, ] ==
  max(raw.rec$ap[, , 1]$values[1, ])) %>%
  as.numeric()

### time of the peak of the APs after its first point
raw.rec$ap[, , 1]$interval * points_to_peak

ap <- raw.rec$ap[, , 1]$times %>% as.double()
ap_peaks <- tibble(peak_times = (ap + c(raw.rec$ap[, , 1]$interval * points_to_peak)))
```



Loading all data calculated by **SyncDesyncAnalysis** function
**SyncDesyncAnalysis** is a function that automatically loads the **.mat** files containing the raw recordings, filters and down samples LFP, finds the AP peak times, calculates ISI and saves everything in data frames. The dataframes are saved in the **sync_desync_tibbles.rds** file


```{r echo=FALSE, message=FALSE, warning=FALSE}
all_eeg_df <- readRDS(file = file.path("data", "sync_desync", "sync_desync_tibbles.rds"))
EEG_ds_df <- purrr::reduce(all_eeg_df, bind_rows)
```

## Plotting

### ISI to determine clusters

```{r echo=FALSE, fig.width=10, fig.height= 10, message=FALSE, warning=FALSE}
gg_isi_periods <- lapply(all_eeg_df, function(x) {
  ggplot(
    data = x %>%
      dplyr::filter(
        AP == 1,
        !is.na(eeg_state),
        state_length >= 1 / first_peak * 2,
        isi > 0, isi < 1 / first_peak
      ), ### NAs in eeg_state because moving_sd has NAs at the end
    mapping = aes(x = isi)
  ) +
    geom_histogram(aes(
      fill = eeg_state,
      color = eeg_state
    ),
    alpha = .3,
    bins = x$nbins %>% unique()
    ) +
    geom_vline(
      xintercept = x$burst_threshold_isi %>% unique()
    ) +
    annotate(
      geom = "text",
      x = x$burst_threshold_isi %>% unique() - x$burst_threshold_isi %>% unique() / 7,
      y = Inf,
      label = paste0("Burst threshold = ", x$burst_threshold_isi %>% unique()),
      angle = 90,
      hjust = 1.2
    ) +
    labs(caption = x$ID %>% unique())
})


### https://github.com/kassambara/ggpubr/issues/158
### bug!!! had to open a NULL pdf device
pdf(NULL)
p_isi <- ggpubr::ggarrange(plotlist = gg_isi_periods, common.legend = TRUE, legend = "bottom")
x <- dev.off()

p_isi
```

### LFP and APs

```{r echo=FALSE, fig.width=10, fig.height=20, message=FALSE, warning=FALSE}
plotting_region <- c(90, 150)

# plotting_region <- c(all_eeg_df[[4]]$times[length(all_eeg_df[[4]]$times)] - 80, all_eeg_df[[4]]$times[length(all_eeg_df[[4]]$times)])

gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}


### rect_index: to show areas excluded from analysis. Based on the length of a period during which clusters can be expected: (1/first_peak) * 2 
eeg_sum_plot <- lapply(all_eeg_df, function(x) {
  rect_index <-
    tibble(
      state_start = subset(x, state_length <= 1 / first_peak * 2) %>%
        dplyr::select(state_start, state_end) %>% pull(state_start) %>% unique(),
      state_end = subset(x, state_length <= 1 / first_peak * 2) %>%
        dplyr::select(state_start, state_end) %>% pull(state_end) %>% unique()
    )

  

  ggplot(
    data = dplyr::filter(x, times > plotting_region[1], times < plotting_region[2]),
    mapping = aes(x = times, y = eeg_values)
  ) +
    geom_line(aes(color = "gray")) +
    geom_line(
      mapping = aes(x = times, y = moving_sd, color = gg_color_hue(2)[1]), size = 1
    ) +
    geom_line(
      mapping = aes(x = times, y = levels + 3, color = gg_color_hue(2)[2]), size = 1
    ) +
    geom_segment(
      mapping = aes(color = "black"),
      x = plotting_region[1] - 2, xend = plotting_region[2] + 2,
      y = x$sd_threshold %>% unique(), yend = x$sd_threshold %>% unique(),
      linetype = "dashed"
    ) +
    geom_point(
      data = . %>% dplyr::filter(!is.na(ap_peak_times)),
      mapping = aes(x = ap_peak_times, y = 5),
      shape = "|",
      size = 4,
      color = "dark gray"
    ) +
    geom_point(
      data = . %>% dplyr::filter(burst_isi == 1),
      mapping = aes(x = times, y = burst_isi + 4),
      color = "black",
      shape = "|",
      size = 5
    ) +
    geom_rect(
      data = dplyr::filter(rect_index, state_start > plotting_region[1], state_start < plotting_region[2]),
      inherit.aes = F,
      mapping = aes(xmin = state_start, xmax = state_end, ymin = -Inf, ymax = +Inf), fill = "pink", alpha = 0.5
    ) +
    ylab("EEG [au]") +
    xlab("Time") +

    ### creating legend
    scale_color_identity(
      name = "Lines",
      guide = guide_legend(override.aes = list(shape = c(rep(NA, 4)))),
      breaks = c("gray", gg_color_hue(2)[1], gg_color_hue(2)[2], "black"),
      labels = c("EEG", "Moving SD", "Levels", "SD threshold")
    ) +
    theme_minimal() +
    labs(caption = x$ID %>% unique())
})

pdf(NULL)
p_eeg <- ggpubr::ggarrange(
  plotlist = eeg_sum_plot,
  common.legend = TRUE,
  legend = "bottom", ncol = 1
)
x <- dev.off()

p_eeg
```

### Number of clusters in sync vs desync periods

Short periods are removed from analysis: (1/first_peak) * 2
**first_peak** is the dominant frequency on the power spectrum, **1/first_peak** is the length of one cycle. Clusters can only be found in the period of two cycles if we assume that the clusters are linked to the cortical oscillation 


```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
left_join(
  EEG_ds_df %>%
    dplyr::filter(!is.na(eeg_state), state_length >= 1 / first_peak * 2) %>%
    dplyr::group_by(ID, eeg_state) %>%
    dplyr::summarise(No_clusters = sum(burst_isi, na.rm = T)),
  EEG_ds_df %>%
    dplyr::filter(!is.na(eeg_state), state_length >= 1 / first_peak * 2) %>%
    dplyr::group_by(ID, eeg_state) %>%
    dplyr::summarise(state_length_sum = sum(state_length %>% unique(), na.rm = T))
) %>%
  dplyr::filter(!grepl(pattern = "01|04", ID)) %>%
  ggplot(aes(x = eeg_state, y = No_clusters / state_length_sum)) +
  geom_point(aes(col = ID), size = 3) +
  geom_line(aes(col = ID, group = ID)) +
  geom_boxplot(
    aes(x = eeg_state, y = No_clusters / state_length_sum),
    inherit.aes = F,
    alpha = 0, width = 0.2, lwd = 1, fatten = 1.2
  ) +
  geom_signif(
    comparisons = list(c("desync", "sync")),
    map_signif_level = TRUE,
    test = "wilcox.test",
    margin_top = .1,
    color = "darkgray"
  ) +
  ggpubr::stat_compare_means(method = "wilcox.test", paired = F)

# geom_text_repel(
#   data = . %>% dplyr::filter(eeg_state == "sync"),
#   aes(label = ID),
#   direction = "y",
#   hjust = -1)
```


<br><br><br><br><br><br>

# Terminals

## Loading data

```{r message=FALSE, warning=FALSE, tidy=TRUE}
file_path <- file.path(
  "data",
  "Emi_terminals",
  "terminal_measures.csv"
)
TERMINALS <- read.table(file = file_path, sep = ";", header = T, na.string = "na")
head(TERMINALS)
```

## Bouton area on different dendritic domains

```{r echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
ggplot(data = TERMINALS, mapping = aes(x = postion, y = b_area)) +
  geom_half_violin(
    position = position_nudge(x = -.15),
    trim = F,
    side = "l",
    scale = "area",
    draw_quantiles = c(0.5),
    aes(fill = postion)
  ) +
  geom_jitter(position = position_jitter(0.1))
```

## Area of boutons with *m2* and *unknown* origin

```{r fig.width=10, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(
  data = TERMINALS,
  mapping = aes(x = origin, y = b_area)
) +
  geom_half_violin(
    position = position_nudge(x = -.15),
    trim = F,
    side = "l",
    scale = "count",
    draw_quantiles = c(0.5)
    # fill = alpha(0.1),
    # aes(col = postion)
  ) +
  geom_jitter(
    position = position_jitter(0.1),
    aes(col = postion)
  ) +
  annotate("text",
    x = 0.7,
    y = 1,
    label = paste(
      "n = ",
      TERMINALS %>%
        dplyr::group_by(origin) %>%
        dplyr::summarise(n = length(b_area)) %>%
        dplyr::filter(origin == "m2") %>%
        select(n) %>%
        as.character()
    )
  ) +
  annotate("text",
    x = 1.6,
    y = 1,
    label = paste(
      "n = ",
      TERMINALS %>%
        dplyr::group_by(origin) %>%
        dplyr::summarise(n = length(b_area)) %>%
        dplyr::filter(origin == "unknown") %>%
        select(n) %>%
        as.character()
    )
  )
```


<br><br><br><br><br><br>

# Methods


## Surgery
Altogether 20 adult male C57Bl/6J BAC GlyT2::eGFP and GlyT2::cre mice (20-30g) were used for the experiments. Surgery, acute recording experiments and implantations were done under ketamine/xylazine anaesthesia. The head of the animal was fixed in a stereotaxic apparatus (David Kopf Instruments, Tujunga, California 91042, Model 900 Small Animal Stereotaxic Instrument). Initially, mice received an intraperitoneal injection of ketamine (111 mg/kg) and xylazine (4.3 mg/kg). For the maintenance of the anaesthesia, intramuscular injection of ketamine/xylazine was given every 30-50 min during the experiments.
 
## In vivo juxtacellular recording and labeling, local field potential (LFP) recording

Bipolar LFP electrodes (FHC, resistance ~1 M) were placed into the frontal cortex of anaesthetised mice (Bregma 1.7 mm; lateral -0.8 mm). Before recording the signal was fed through a signal pre-processing pipeline: amplification and band-pass filtering from 0.16 Hz to 5 kHz and from 100 Hz to 5 kHz to record the fast multiunit activity (Supertech BioAmp, Supertech, Pcs, Hungary). The signal was digitized at 20 kHz (micro 1401 mkii, CED, Cambridge, UK). 

To test the direct mono synaptic connection between the IL and the recorded PRF neurons concentric bipolar stimulating electrodes were inserted into the IL (Bregma -1.5 mm; lateral 2 mm; ventral from cortical surface -3.2 mm, at 20 degrees, electrode separation 0.8mm). PRF single unit activity was recorded by glass microelectrodes (in vivo impedance of 20-40 M) made from borosilicate glass capillaries (1.5 mm outer diameter, 0.75 or 0.86 inner diameter, Sutter Instrument Co., Novato, CA, USA or WPI Inc. Sarasota, Fl, USA) and filled with 0.5 M K+- acetate and 2% neurobiotin (Vector Laboratories, Burlingame, CA, USA). Electrodes were positioned into the PRF (Bregma -4.4 mm; lateral -0.8 to -1 mm; ventral from cortical surface -3.8 to -4.8 mm) using a piezoelectric microdrive (Burleigh 6000 ULN or ISS 8200, EXFO, Quebec City, Quebec, Canada). Neuronal signals were amplified by a DC amplifier (Axoclamp 2B, Axon Instruments/Molecular Devices, Sunnyvale, CA, USA), further amplified and filtered between 0.16 Hz and 5 kHz by a signal conditioner (LinearAmp, Supertech) and recorded by Spike2 5.0 (CED). Juxtacellular labeling of the recorded neurons was done as described previously (Pinault, 1996). Briefly, after recording the activity of the neuron positive current steps (0.5-8 nA) were applied at 2 Hz via the neurobiotin filled recording pipette to induce firing. The neuron was active only during these current steps and took up neurobiotin, which filled the soma and proximal dendrites. In some cases, distal dendritic regions and axons were also labeled.

After the the experiment the animal was perfused and coronal sections (50 m) were cut from the PRF. The neurons were visualized using Cy3 conjugated streptavidin (1:2,000, Jackson) and GlyT2::eGFP positivity was determined by confocal microscopy. The neurons were then developed using ABC and DAB-Ni, and the sections containing the labeled neurons were dehydrated and embedded in Durcupan. The dendritic trees of the labeled neurons were reconstructed using Neurolucida 5.2 software (MBF Bioscience, Magdeburg, Germany).

## Analysis


All analysis was done in R Statistical Computing Software (R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/). For the use packages see the "Used packeges" section.

### Sync-desync analysis



n=7 recordings from n=? animals
Raw LFP signal was filtered, down-sampled and standardized. The power spectral density of each LFP signal was calculated and the dominant frequency was determined. To find the synchronous and desynchronous periods of the cortical LFP signal the standard deviation of the signal amplitude was calculated in overlapping sliding windows. The size of the sliding window was equal to the length of one cycle the overlap was half a cycle. The threshold for synchronous and desynchronous periods was placed half way between the median and the 25th percentile.  

Action potential clusters were determined by separating the two peaks at the minimum of the bimodal inter-spike interval histograms (ISI). Number of bins of the ISI was determined by the Freedman-Diaconis rule for optimal bin-width, the window size was equal to the length of one cycle (period) of the cortical slow oscillation.


Determine if a cell fires in clusters or not: dip test.



### Used R packeges:


**tidyverse** 
"Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686"

**R.matlab**
"Henrik Bengtsson (2018). R.matlab: Read and Write MAT Files and Call MATLAB from Within R. R package version 3.6.2. https://CRAN.R-project.org/package=R.matlab"

**reshape**
"Hadley Wickham (2007). Reshaping Data with the reshape Package. Journal of Statistical Software, 21(12), 1-20. URL http://www.jstatsoft.org/v21/i12/."

**bspec**
"Christian Roever (2015). bspec: Bayesian spectral inference. R package version 1.5."


"Roever, C., Meyer, R. and Christensen, N. (2010). Modelling coloured residual noise in gravitational-wave signal processing. Classical and Quantum Gravity, 28(1):015010. See also Arxiv preprint 0804.3853, URL http://arxiv.org/abs/0804.3853."


**WaveletComp**
"Angi Roesch and Harald Schmidbauer (2018). WaveletComp: Computational Wavelet Analysis. R package version 1.1. https://CRAN.R-project.org/package=WaveletComp"

**diptest**
"Martin Maechler (2021). diptest: Hartigan's Dip Test Statistic for Unimodality - Corrected. R package version 0.76-0. https://CRAN.R-project.org/package=diptest"

**ggsignif**
"Ahlmann-Eltze, C., & Patil, I. (2021). ggsignif: R Package for Displaying Significance Brackets for 'ggplot2'. PsyArxiv. doi:10.31234/osf.io/7awm6"

**ggrepel**
"Kamil Slowikowski (2021). ggrepel: Automatically Position Non-Overlapping Text Labels with 'ggplot2'. R package version 0.9.1. https://CRAN.R-project.org/package=ggrepel"

**gghalves**
"Frederik Tiedemann (2020). gghalves: Compose Half-Half Plots Using Your Favourite Geoms. R package version 0.1.1. https://CRAN.R-project.org/package=gghalves"


"Ldecke, Patil, Ben-Shachar, Wiernik, Waggoner & Makowski (2020). Visualisation Toolbox for 'easystats' and Extra Geoms, Themes and Color Palettes for 'ggplot2'. CRAN. Available from https://easystats.github.io/see/"

**DT**
"Yihui Xie, Joe Cheng and Xianying Tan (2021). DT: A Wrapper of the JavaScript Library 'DataTables'. R package version 0.18. https://CRAN.R-project.org/package=DT"

**ggpubr**
"Alboukadel Kassambara (2020). ggpubr: 'ggplot2' Based Publication Ready Plots. R package version 0.4.0. https://CRAN.R-project.org/package=ggpubr"

**styler**
"Kirill Mller and Lorenz Walthert (2021). styler: Non-Invasive Pretty Printing of R Code. R package version 1.5.1. https://CRAN.R-project.org/package=styler"

**formatR**
"Yihui Xie (2021). formatR: Format R Code Automatically. R package version 1.11. https://CRAN.R-project.org/package=formatR"

**exactRankTests**
"Torsten Hothorn and Kurt Hornik (2021). exactRankTests: Exact Distributions for Rank and Permutation Tests. R package version 0.8-32. https://CRAN.R-project.org/package=exactRankTests"

**signal**
"signal developers (2013). signal: Signal processing. URL: http://r-forge.r-project.org/projects/signal/."

**STAR**
"Christophe Pouzat (2012). STAR: Spike Train Analysis with R. R package version 0.3-7. https://CRAN.R-project.org/package=STAR"

**seewave**
"Sueur J, Aubin T, Simonis C (2008). seewave: a free modular tool for sound analysis and synthesis. Bioacoustics, 18: 213-226"

**mixtools**
"Tatiana Benaglia, Didier Chauveau, David R. Hunter, Derek Young (2009). mixtools: An R Package for Analyzing Finite Mixture Models. Journal of Statistical Software, 32(6), 1-29. URL http://www.jstatsoft.org/v32/i06/."

**RcppRoll**
"Kevin Ushey (2018). RcppRoll: Efficient Rolling / Windowed Operations. R package version 0.3.0. https://CRAN.R-project.org/package=RcppRoll"

**gganimate**
"Thomas Lin Pedersen and David Robinson (2020). gganimate: A Grammar of Animated Graphics. R package version 1.0.7. https://CRAN.R-project.org/package=gganimate"

**transformr**
"Thomas Lin Pedersen (2020). transformr: Polygon and Path Transformations. R package version 0.1.3. https://CRAN.R-project.org/package=transformr"

**gifski**
"Jeroen Ooms (2021). gifski: Highest Quality GIF Encoder. R package version 1.4.3-1. https://CRAN.R-project.org/package=gifski"

**matlib**
"Michael Friendly, John Fox and Phil Chalmers (2020). matlib: Matrix Functions for Teaching and Learning Linear Algebra and nMultivariate Statistics. R package version 0.9.4. https://CRAN.R-project.org/package=matlib"


