---
title: "Halo freely moving analysis"
author: "Viktor Plattner"
output:
  html_document:
    css: styles.css
    number_sections: yes
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: true
spacing: single
fontsize: 12pt
always_allow_html: yes
header-includes:
   - \usepackage{animate}

params:
  SELECTED_ANIMAL: "FM9"
  SELECTED_ANIMAL_GRP: "FM1_right"


knit: (
  function (inputFile, encoding) {
    
    rmarkdown::render(  
      input = inputFile,
      encoding = encoding,
      output_file = file.path(
        "notebook_html", 
        paste0("freely_moving_analysis_", 
        rmarkdown::yaml_front_matter(inputFile)$params$SELECTED_ANIMAL,
        "_",
        Sys.Date(),
         ".html")
        ),
      envir = globalenv()
      )
    }
---


```{r setup, include=FALSE}
source(file.path("..","supplementary_functions","md_analysis_setup_script.R"))
# Brief explanation of the sourced script:
# - Loads `tidyverse` for data manipulation and visualization.
# - Defines custom operators/functions for enhanced R capabilities.
# - Dynamically loads additional libraries based on patterns in `.Rmd`/.R` files.
# - Sets a global theme for ggplot objects.
# - Includes custom functions for data filtering and recalculating distances/speeds.

# Define global constants and settings
SCALE <- 4 # pixel/mm
FPS <- 30 # frame rate
STANDARD_STIM_LENGTH <- 10 # Default stimulus length in seconds
AROUND_STIM_LENGTH <- FPS * 8 # Stimulus length around default
TIME_RES <- 1 / FPS # Time between frames
ANG_THRESHOLD <- 90 # Angle threshold
MOUSE_SPEED_MPS <- 3.3 # Mouse speed in meters per second
DATAPOINT_TO_PLOT <- 6 # Data points to plot
XLIM_MM <- c(0, 1280 / SCALE) # X-axis limits
YLIM_MM <- c(0, 720 / SCALE) # Y-axis limits
SELECTED_ANIMAL <- params$SELECTED_ANIMAL
SELECTED_ANIMAL_GRP <- params$SELECTED_ANIMAL_GRP
SELECTED_INTENSITIES <- c("1 mW", "5 mW", "10 mW")
CONTROL_ANIMALS <- c("")
DOC_TITLE <- "Halo freely moving analysis"
```


# Loading and orginising data

```{r warning=F, message=F, echo=FALSE}
stims_halo <- read.csv(
  file.path("data", "PnO_stim_halo", "pno_halo_stims.csv")
) %>%
  as_tibble() %>%
  mutate(
    intensity = paste0(intensity, " mW"),
    stim_length = end - start,
    animal = animal %>% str_sub(-3,-1),
    animal_grp = paste0(animal, "_", stim_side)
  ) %>%
  select(-stim_duration) %>%
  rename("stim_number" = stim.number) %>%
  gather(key = "timing", value = "frame", start, end)

```



```{r message=F, echo = FALSE, warning = FALSE}
coords_raw_halo <- read.csv(
  file.path("data", "PnO_stim_halo", "pno_halo.csv")
) %>%
  as_tibble() %>%
  mutate(
    animal = session %>% substr(1, 3) %>% str_to_upper(),
    animal_grp = case_when(
      session %>% str_detect("bal") ~ paste0(animal, "_left"),
      session %>% str_detect("jobb") ~ paste0(animal, "_right"),
      TRUE ~ NA_character_ # Default value if neither condition is met
    )
  ) %>% 
  full_join(stims_halo) %>% 
  ### creating unique session IDs
  group_by(animal_grp) %>%
  mutate(
    s2 = lag(session, default = "1"),
    s3 = ifelse(s2 == session, 0, 1),
    session = cumsum(s3),
    session = paste0("session_", session)
  ) %>%
  select(-s2, -s3) %>%
  ungroup() %>%
  dplyr::group_by(animal_grp, session) %>%
  dplyr::mutate(
    stim_number = ifelse(timing == "end", NA, stim_number)
  ) %>%
  mutate(grp = cumsum(!is.na(stim_number))) %>%
  dplyr::group_by(animal_grp, session, grp) %>%
  fill(timing) %>%
  mutate(
    stim_freq = replace(stim_freq, timing == "start", stim_freq[1]),
    stim_freq = replace(stim_freq, timing == "end", NA),
    stim_length = replace(stim_length, timing == "start", stim_length[1]),
    stim_length = replace(stim_length, timing == "end", NA),
    intensity = replace(intensity, timing == "start", intensity[1]),
    intensity = replace(intensity, timing == "end", NA),
    stim_number = replace(stim_number, timing == "start", stim_number[1]),
    stim_number = replace(stim_number, timing == "end", NA),
    stim_side = replace(stim_side, timing == "start", stim_side[1]),
    stim_side = replace(stim_side, timing == "end", NA),
    quality = replace(quality, timing == "start", quality[1]),
    quality = replace(quality, timing == "end", NA),
    stim = F,
    stim = replace(stim, timing == "start", T)
  ) %>%
  ungroup() %>%
  group_by(animal_grp, session) %>%
  dplyr::mutate(
    rec_time = (frame / FPS) %>% round(digits = 2),
    stim_categ = case_when(
      stim == T ~ "stim",
      lead(stim, n = AROUND_STIM_LENGTH) == T ~ "before stim",
      lag(stim, n = AROUND_STIM_LENGTH) == T ~ "after stim",
      stim == F ~ "no stim"
    ),

    # shift values to the beginning of the stimulus block ...
    # ... to be able to fill missing values in the next step
    stim_block = ifelse(stim_categ == "no stim", F, T),
    stim_number = lead(stim_number, n = AROUND_STIM_LENGTH),
    intensity = lead(intensity, n = AROUND_STIM_LENGTH),
    stim_freq = lead(stim_freq, n = AROUND_STIM_LENGTH),
    stim_side = lead(stim_side, n = AROUND_STIM_LENGTH),
    quality = lead(quality, n = AROUND_STIM_LENGTH)
  ) %>%
  ungroup() %>%
  group_by(animal_grp, session, stim_block) %>%
  fill(stim_number) %>%
  fill(intensity) %>%
  fill(stim_freq) %>%
  fill(stim_side) %>%
  fill(quality) %>%
  # length of the stimulus
  unite(col = "head", head1, head2, sep = "_") %>%
  unite(col = "center", center1, center2, sep = "_") %>%
  unite(col = "tail", tail1, tail2, sep = "_") %>%
  gather(
    key = "body_part",
    value = "coordinate",
    head, center, tail
  ) %>%
  tidyr::extract(
    col = coordinate, into = c("X", "Y"),
    regex = "([[:alnum:]]+)_([[:alnum:]]+)"
  ) %>%
  dplyr::mutate(
    X = replace(X, miss == 0, NA),
    Y = replace(Y, miss == 0, NA)
  ) %>%
  dplyr::mutate(
    stim_number = replace(stim_number, is.na(stim_number), "no stim")
  )
```


```{r}
coords_halo <- coords_raw_halo %>%
  dplyr::filter(
    body_part != "tail",
  ) %>%
  # this block calculates distances and speed across frames
  dplyr::group_by(animal_grp, body_part, session) %>%
  dplyr::mutate(

    # interpolating missing values
    X_interp = imputeTS::na_interpolation(X %>% as.integer()) %>% round(),
    Y_interp = imputeTS::na_interpolation(Y %>% as.integer()) %>% round(),
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),

    # sometimes the tracking software looses the point and jumps...
    # ...back and forth between unrealistic numbers...
    # ...or starts to follow something else
    # these points are replaced by 0...
    # ...(the animal did not change position between those frames)
    # outliers are defined as Q1 - 3*IQR, Q3 + 3*IQR

    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / SCALE,
    speed_instant = d_mm / TIME_RES,
    mmps = speed_instant * (1 / TIME_RES),
    mps = (speed_instant * (1 / TIME_RES)) / 1000,
    sec_counter = cumsum(ifelse(rec_time %% 1 == 0, 1, 0))
  ) %>%
  group_by(animal_grp, session, body_part, sec_counter) %>%
  dplyr::mutate(speed_sec = sum(d_mm, na.rm = T)) %>%
  ungroup() %>%
  # removing "bad" quality stimuli (repeats)
  dplyr::filter(
    quality != "bad" | is.na(quality),
    stim_number %in% c("1", "2", "3", "4", "5", "no stim")
  ) %>%
  ungroup() %>%
  dplyr::group_by(animal_grp, body_part, session) %>%
  # adding the number of repeats in each session, (-1 to remove "no stim")
  dplyr::mutate(
    No_repeats = stim_number %>% unique() %>% length() - 1
  ) %>%
  ungroup() %>%
  # adding the number of repeats in each intensity group
  # (no -1 as "intensity" variable doesn't run through the session)
  dplyr::group_by(animal_grp, body_part, intensity) %>%
  dplyr::mutate(
    No_repeats_intensity = stim_number %>% unique() %>% length()
  )



coords_halo %>%
  dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP) %>%
  dplyr::group_by(session, body_part) %>%
  slice(1:5)
```



```{r}
coords_halo_down_sapmpled <- coords_halo %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  # this block calculates distances and speed across frames
  dplyr::group_by(animal_grp, body_part, session) %>%
  dplyr::mutate(

    # interpolating missing values
    X_interp = imputeTS::na_interpolation(X %>% as.integer()) %>% round(),
    Y_interp = imputeTS::na_interpolation(Y %>% as.integer()) %>% round(),
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),

    # sometimes the tracking software looses the point and jumps...
    # ...back and forth between unrealistic numbers...
    # ...or starts to follow something else
    # these points are replaced by 0...
    # ...(the animal did not change position between those frames)
    # outliers are defined as Q1 - 3*IQR, Q3 + 3*IQR

    X_diff = ifelse(
      test = X_diff < quantile(
        X_diff,
        na.rm = T
      )["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(
          X_diff,
          na.rm = T
        )["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(
        Y_diff,
        na.rm = T
      )["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(
          Y_diff,
          na.rm = T
        )["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / SCALE,
    speed_instant = d_mm / TIME_RES,
    mmps = speed_instant * (1 / TIME_RES),
    mps = (speed_instant * (1 / TIME_RES)) / 1000,
    sec_counter = cumsum(ifelse(rec_time %% 1 == 0, 1, 0))
  )

coords_halo_down_sapmpled %>%
  dplyr::filter(animal == SELECTED_ANIMAL) %>%
  dplyr::group_by(session, body_part) %>% 
  slice(1:5)
```


```{r message=F, warning=F, echo=F, include=T}
lengths <- coords_halo %>%
  dplyr::filter(
    body_part == "head",
    stim_categ != "no stim",
  ) %>%
  group_by(animal_grp, session, stim_number, intensity, quality) %>%
  count(stim_categ) %>%
  spread(stim_categ, n) %>%
  select(
    animal_grp, session, stim_number,
    intensity, quality, `before stim`,
    stim, `after stim`
  )


n_repeat <- lengths %>%
  dplyr::group_by(animal_grp, intensity) %>%
  dplyr::filter(intensity %in% SELECTED_INTENSITIES) %>%
  summarise(length(stim_number))

lengths
```



```{r}
text_layer <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP, # selected_control
    stim == T
  ) %>%
  group_by(session) %>%
  summarise(int_label = intensity %>% unique())

```

<br> Creating plot colors <br>

```{r}
gg_color_hue <- function(n) {
  hues <- seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n <- 6
cols <- gg_color_hue(n)
```


# Analysis (selected animal: `r SELECTED_ANIMAL`)

## Checking tracking quality

```{r }
plt_xpos <- coords_halo %>%
  dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP, body_part != "tail") %>%
  ggplot(mapping = aes(x = rec_time, y = X_interp %>% as.integer())) +
  geom_line(aes(col = body_part)) +
  theme(legend.position = "none") +
  ylab("X position") +
  facet_wrap(~session, ncol = 1) +
  ggtitle(paste0("X position: ", SELECTED_ANIMAL))



plt_ypos <- coords_halo %>%
  dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP, body_part != "tail") %>%
  ggplot(mapping = aes(x = rec_time, y = Y %>% as.integer())) +
  geom_line(aes(col = body_part)) +
  theme(legend.position = "none") +
  ylab("Y position") +
  facet_wrap(~session, ncol = 1) +
  ggtitle(paste0("Y position: ", SELECTED_ANIMAL))

cowplot::plot_grid(plt_xpos, plt_ypos, ncol = 2)
```



```{r }
coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part != "tail"
  ) %>%
  dplyr::group_by(session, frame) %>%
  dplyr::mutate(
    dx = diff(X_interp),
    dy = diff(Y_interp),
    # ldx = lead(dx),
    # ldy = lead(dy),
    hc_d = sqrt(dx^2 + dy^2)
  ) %>%
  ungroup() %>%
  dplyr::filter(body_part == "center") %>%
  ggplot(
    mapping = aes(x = rec_time, y = hc_d)
  ) +
  geom_hline(
    aes(
      yintercept = quantile(
        hc_d,
        na.rm = T
      )["25%"] - 1.5 * IQR(hc_d, na.rm = T)
    ),
    col = "gray50",
    lty = "dashed"
  ) +
  geom_hline(
    aes(yintercept = quantile(
      hc_d,
      na.rm = T
    )["75%"] + 1.5 * IQR(hc_d, na.rm = T)),
    col = "gray50",
    lty = "dashed"
  ) +
  geom_line(
    aes(col = stim, group = 1)
  ) +
  geom_label_repel(
    data = . %>%
      dplyr::filter(stim_number != "no stim", stim_categ == "stim") %>%
      dplyr::group_by(session, stim_number) %>%
      slice(1),
    aes(
      x = rec_time,
      y = mean(hc_d, na.rm = T),
      label = quality
    ),
    direction = "x",
    nudge_x = 1
  ) +
  facet_wrap(~session, ncol = 1, scales = "free_y") +
  coord_cartesian(
    ylim = c(20, 130)
  ) +
  global_facet_theme
```


## Plotting trajectories

```{r echo=, fig.height=10, fig.width=15, warning=FALSE}
plt_traject_stim <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    # intensity %in% SELECTED_INTENSITIES,
    stim == T,
    frame %% 5 == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / SCALE,
      y = Y_interp, # / SCALE,
      label = frame
    )
  ) +

  ### drawing positions during no stimulus
  ### "coords_halo" was filtered differently compared to the stimulus dataset (number of frames)
  geom_point(
    data = coords_halo %>%
      dplyr::filter(
        animal_grp == SELECTED_ANIMAL_GRP,
        # intensity %in% c(SELECTED_INTENSITIES, NA),
        body_part == "center",
        stim == F,
        frame %% 5 == 0
      ),
    size = 1,
    alpha = 0.1
  ) +

  ### drawing positions during stimulus
  ### body parts are connected by lines
  # geom_line(
  #   aes(group = frame, col = as.character(stim_number)),
  # ) +
  geom_path(
    aes(group = frame, col = as.character(stim_number)),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm")
    )
  ) +
  # geom_point(
  #   aes(shape = body_part, col = as.character(stim_number)),
  #   size = 2
  # ) +
  xlim(c(0, 1280)) + # / SCALE)) +
  ylim(c(0, 720)) + # / SCALE)) +
  coord_fixed() +
  facet_wrap(~session, nrow = 1) +
  geom_label(
    data = text_layer %>% dplyr::filter(int_label %in% SELECTED_INTENSITIES),
    aes(
      x = Inf, y = -Inf,
      hjust = 1.1, vjust = -0.5,
      label = paste0("Intensity: ", int_label)
    )
  ) +
  labs(col = "Stimulus number", shape = "Body part") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(SELECTED_ANIMAL_GRP, " trajectory")) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom"
  )



```

```{r}
cowplot::plot_grid(plt_traject_stim)
```

```{r}
plt_taject_stim_baseline <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    intensity == "5 mW",
    stim == F,
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / SCALE,
      y = Y_interp, # / SCALE,
      label = frame
    )
  ) +
  geom_path(
    aes(group = frame, col = as.character(stim_number)),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm")
    )
  ) +
  xlim(c(0, 1280)) + # / SCALE)) +
  ylim(c(0, 720)) + # / SCALE)) +
  coord_fixed() +
  # geom_label(
  #   data = text_layer %>% dplyr::filter(int_label %in% SELECTED_INTENSITIES),
  #   aes(
  #     x = Inf, y = -Inf,
  #     hjust = 1.1, vjust = -0.5,
  #     label = paste0("Intensity: ", int_label))) +
  labs(col = "Stimulus number") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(SELECTED_ANIMAL_GRP, " baseline trajectory")) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  )


plt_taject_stim_stim <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    intensity == "5 mW",
    stim == T,
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / SCALE,
      y = Y_interp, # / SCALE,
      label = frame
    )
  ) +
  geom_path(
    aes(group = frame, col = as.character(stim_number)),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm")
    )
  ) +
  xlim(c(0, 1280)) + # / SCALE)) +
  ylim(c(0, 720)) + # / SCALE)) +
  coord_fixed() +
  # geom_label(
  #   data = text_layer %>% dplyr::filter(int_label %in% SELECTED_INTENSITIES),
  #   aes(
  #     x = Inf, y = -Inf,
  #     hjust = 1.1, vjust = -0.5,
  #     label = paste0("Intensity: ", int_label))) +
  labs(col = "Stimulus number") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(SELECTED_ANIMAL, " stimulus trajectory")) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  )



#### with facet wrap

plt_taject_stim_wrap <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    intensity == "5 mW",
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ggplot(
    mapping = aes(
      x = X_interp, # / SCALE,
      y = Y_interp, # / SCALE,
      label = frame
    )
  ) +
  geom_path(
    aes(group = frame),
    arrow = grid::arrow(
      ends = "first",
      type = "closed",
      angle = 10,
      length = unit(3, "mm")
    )
  ) +
  xlim(c(0, 1280)) + # / SCALE)) +
  ylim(c(0, 720)) + # / SCALE)) +
  coord_fixed() +
  # geom_label(
  #   data = text_layer %>% dplyr::filter(int_label %in% SELECTED_INTENSITIES),
  #   aes(
  #     x = Inf, y = -Inf,
  #     hjust = 1.1, vjust = -0.5,
  #     label = paste0("Intensity: ", int_label))) +
  labs(col = "Stimulus number") +
  xlab("X position [pixel]") +
  ylab("Y position [pixel]") +
  ggtitle(paste0(SELECTED_ANIMAL, " trajectory")) +
  facet_wrap(~stim) +
  global_facet_theme +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  )

```

Baseline trajectory: only before and after the stimulus, not the whole session

```{r}
cowplot::plot_grid(plt_taject_stim_baseline, plt_taject_stim_stim, nrow = 1)
```


## Recalculating distances between tracked points after leaving out frames

```{r echo = FALSE}
plt_xlim <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    # frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  dplyr::slice(1:11) %>%
  pull(X_interp) %>%
  range() + c(-1, 1)

plt_ylim <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    # frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  dplyr::slice(1:11) %>%
  pull(Y_interp) %>%
  range() + c(-1, 1)


plt1_path <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    session == "session_1"
  ) %>%
  dplyr::slice(1:6) %>%
  dplyr::arrange(frame) %>% ### to be used with geom_path
  ggplot(
    mapping = aes(x = X_interp, y = Y_interp)
  ) +
  geom_text(
    aes(
      color = frame %>% as.factor(),
      label = frame
    ),
    size = 6
  ) +
  geom_segment(
    aes(
      x = X_interp[1], xend = X_interp[2],
      y = Y_interp[1], yend = Y_interp[1]
    ),
    linetype = "dashed", col = "red"
  ) +
  geom_segment(
    aes(
      x = X_interp[2], xend = X_interp[2],
      y = Y_interp[1], yend = Y_interp[2]
    ),
    linetype = "dashed", col = "red"
  ) +
  geom_path(col = "gray", linetype = "dashed") +
  geom_label_repel(
    data = . %>% dplyr::slice(1:5),
    aes(
      x = X_interp + X_diff / 2,
      y = Y_interp + Y_diff / 2,
      label = d %>% round(4)
    ),
    direction = "x",
    nudge_x = -.7
  ) +
  theme(legend.position = "none") +
  xlim(plt_xlim) +
  ylim(plt_ylim) +
  coord_fixed(ratio = 1)


plt2_path <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    session == "session_1",
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ### recalculates distances after frames were removed
  dplyr::group_by(session) %>%
  dplyr::mutate(
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / SCALE,
    speed_instant = d_mm / TIME_RES,
    mmps = speed_instant * (1 / TIME_RES),
    mps = (speed_instant * (1 / TIME_RES)) / 1000
  ) %>%
  dplyr::slice(1:2) %>%
  dplyr::arrange(frame) %>% ### to be used with geom_path
  ggplot(
    mapping = aes(x = X_interp, y = Y_interp)
  ) +
  geom_text(
    aes(
      color = frame %>% as.factor(),
      label = frame
    ),
    size = 6
  ) +
  geom_segment(
    aes(
      x = X_interp[1], xend = X_interp[2],
      y = Y_interp[1], yend = Y_interp[1]
    ),
    linetype = "dashed", col = "red"
  ) +
  geom_segment(
    aes(
      x = X_interp[2], xend = X_interp[2],
      y = Y_interp[1], yend = Y_interp[2]
    ),
    linetype = "dashed", col = "red"
  ) +
  geom_path(col = "gray", linetype = "dashed") +
  geom_label_repel(
    data = . %>% dplyr::slice(1:1),
    aes(
      x = X_interp + X_diff / 2,
      y = Y_interp + Y_diff / 2,
      label = d %>% round(4)
    ),
    direction = "x",
    nudge_x = -.7
  ) +
  theme(legend.position = "none") +
  xlim(plt_xlim) +
  ylim(plt_ylim) +
  coord_fixed(ratio = 1)

library(cowplot)

cowplot::plot_grid(plt1_path, plt2_path, ncol = 1)
```

## Rolling mean of traveled distance

Plotting the mean traveled distance calculated in a rolling window (1 sec). The mean was calculated separately during stimulus (rolling window restarts with the beginning of stimuli). Colors represent different periods of the recording (no-stimulus, before stimulus, during stimulus, after stimulus).

Defining line colors

```{r}
line_cols <- gg_color_hue(3)

group_colors <- c(
  "no stim" = "gray30", "before stim" = line_cols[1],
  "stim" = line_cols[3], "after stim" = line_cols[2]
)
group_colors2 <- c(
  "no stim" = "gray30", "right before" = line_cols[1],
  "stim" = line_cols[3], "right after" = line_cols[2]
)
```

### Example session

```{r}
coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    session == "session_3",
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ### recalculates distances after frames were removed
  dplyr::group_by(session) %>%
  dplyr::mutate(
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / SCALE,
    speed_instant = d_mm / TIME_RES,
    mmps = speed_instant * (1 / TIME_RES),
    mps = (speed_instant * (1 / TIME_RES)) / 1000
  ) %>%
  dplyr::group_by(stim) %>%
  dplyr::mutate(mean_mps_roll = zoo::rollmean(x = d_mm, k = FPS, fill = NA)) %>%
  ### plotting
  ggplot(aes(x = rec_time, y = d_mm)) +
  geom_point(aes(alpha = stim)) +
  geom_line(aes(x = rec_time, y = mean_mps_roll, col = stim_categ, group = 1), size = .8) +
  ggtitle(paste0(SELECTED_ANIMAL_GRP, " mean traveled distance by stimulus period ", "(center)")) +
  scale_color_manual(values = group_colors)
```

### Center

```{r}
coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    body_part == "center",
    frame %% DATAPOINT_TO_PLOT == 0
  ) %>%
  ### recalculates distances after frames were removed
  dplyr::group_by(session) %>%
  dplyr::mutate(
    X_diff = c(diff(X_interp), NA) %>% round(4),
    Y_diff = c(diff(Y_interp), NA) %>% round(4),
    X_diff = ifelse(
      test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
        X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
      yes = 0,
      no = X_diff
    ),
    Y_diff = ifelse(
      Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
        Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
      yes = 0,
      no = Y_diff
    ),
    d = sqrt(X_diff^2 + Y_diff^2),
    d_mm = d / SCALE,
    speed_instant = d_mm / TIME_RES,
    mmps = speed_instant * (1 / TIME_RES),
    mps = (speed_instant * (1 / TIME_RES)) / 1000
  ) %>%
  dplyr::group_by(session, stim_categ) %>%
  dplyr::mutate(mean_mps_roll = zoo::rollmean(x = d_mm, k = FPS, fill = NA)) %>%
  ### plotting
  ggplot(aes(x = rec_time, y = d_mm)) +
  geom_point(aes(alpha = stim)) +
  geom_line(aes(x = rec_time, y = mean_mps_roll, col = stim_categ, group = 1), size = 1) +
  facet_wrap(~session, scales = "free", ncol = 1) +
  geom_label(data = text_layer, aes(x = -Inf, y = Inf, hjust = -.2, vjust = 1.3, label = paste0("Intensity: ", int_label))) +
  ggtitle(paste0(SELECTED_ANIMAL_GRP, " mean traveled distance by stimulus period ", "(center)")) +
  scale_color_manual(values = group_colors) +
  global_facet_theme
```

## Sum traveled distance by individual stimuli within session (1s before and during the stimulus)

Plotting traveled distance during individual stimuli (4 to 6 repeats) within sessions. Traveled distance is summed through the last second of "before stimulus" category and through first second of the stimulus. Traveled distance "after stimulus" is not included. <br>

The same data is shown on the plots with different groupings: <br>

-   intensities are grouped together (colors), stimulus numbers are separated (panels) <br>
-   stimulus numbers are grouped together (colors), intensities are separated (panels) <br>

<br> The table shows the filtered data (1s before and 1s during stimulus) <br>

```{r message=F, warning=F, echo=F}
coords_halo_1s_window <- bind_rows(
  ### before stim
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "before stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
    dplyr::group_by(animal_grp, body_part, intensity, stim_side, stim_categ, stim_number) %>%
    dplyr::slice(tail(row_number(), 6)),
  ### stim
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
    dplyr::group_by(animal_grp, body_part, intensity, stim_categ, stim_number) %>%
    dplyr::slice(1:6)
) %>% arrange(animal_grp, body_part, session, stim_number, stim_categ)


coords_halo_1s_window %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    intensity %in% SELECTED_INTENSITIES
  ) %>%
  select(session, animal_grp, frame, rec_time, stim_categ, d, stim_number, intensity, stim_side, body_part, X_diff, Y_diff) %>% 
  head(1000)
```

```{r message=F, warning=F, echo=F}
coords_halo_8s_window <- bind_rows(
  ### before stim
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "before stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
    dplyr::group_by(animal_grp, body_part, intensity, stim_side, stim_categ, stim_number) %>%
    dplyr::slice(tail(row_number(), 41)),
  ### stim
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "stim"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
    dplyr::group_by(animal_grp, body_part, intensity, stim_categ, stim_number) %>%
    dplyr::slice(1:41)
) %>% arrange(animal_grp, body_part, session, stim_number, stim_categ)


coords_halo_8s_window %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    intensity %in% SELECTED_INTENSITIES
  ) %>%
  select(session, animal_grp, frame, rec_time, stim_categ, d, stim_number, intensity, stim_side, body_part, X_diff, Y_diff) %>%
  head(200)
```


## Calculating angles

### First 10 head - center vectors

```{r echo = FALSE ,message=FALSE, warning=FALSE}
coords_halo %>%
  dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP) %>%
  dplyr::group_by(body_part) %>%
  slice(1:10) %>%
  dplyr::filter(body_part != "tail") %>%
  ggplot(
    mapping = aes(x = X, y = Y)
  ) +
  geom_point(aes(shape = body_part), size = 2) +
  geom_line(aes(group = frame, color = as.character(frame))) +
  geom_text_repel(aes(label = paste0(X, ",", Y), col = as.character(frame))) +
  geom_segment(
    data = coords_halo %>%
      dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP) %>%
      dplyr::group_by(body_part) %>%
      slice(1:10) %>%
      dplyr::filter(body_part != "tail", frame == 0),
    aes(
      x = X[1], xend = X[2],
      y = Y[1], yend = Y[1]
    ), linetype = "dashed", col = "red"
  ) +
  geom_segment(
    data = coords_halo %>%
      dplyr::filter(animal_grp == SELECTED_ANIMAL_GRP) %>%
      dplyr::group_by(body_part) %>%
      slice(1:10) %>%
      dplyr::filter(body_part != "tail", frame == 0),
    aes(
      x = X[2], xend = X[2],
      y = Y[1], yend = Y[2]
    ), linetype = "dashed", col = "red"
  )
```

```{r eval=FALSE}
coords_halo %>%
  dplyr::filter(animal == "FM1") %>%
  dplyr::group_by(body_part) %>%
  slice(1:3) %>%
  dplyr::filter(body_part != "tail", frame == 0) %>%
  pull(X)
```


### Calculating signed angles

```{r message=F, warning=F, echo = FALSE}
corrected_angles <- coords_halo %>%
  dplyr::filter(
    animal_grp == SELECTED_ANIMAL_GRP,
    # frame %% DATAPOINT_TO_PLOT == 0, ### here it's good to have all the data points to calculate the cumulative angles
    body_part != "tail"
  ) %>%
  dplyr::group_by(session, frame) %>%
  mutate(
    vectx = diff(X_interp),
    vecty = diff(Y_interp)
  ) %>%
  dplyr::group_by(session) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>%
  mutate(
    signed_angle_corrected = case_when(
      signed_angle >= 180 ~ 360 - signed_angle,
      signed_angle %>% abs() < 180 ~ signed_angle,
      signed_angle <= -180 ~ -360 - signed_angle,
      signed_angle > -180 | signed_angle < 0 ~ signed_angle
    )
  ) %>%
  ### removing unrealistically large angles
  mutate(signed_angle_corrected = ifelse(signed_angle_corrected %>% abs() > ANG_THRESHOLD, 0, signed_angle_corrected))
```


### Plotting cumulative angles

```{r}
corrected_angles %>%
  dplyr::filter(body_part == "center") %>%
  group_by(session) %>%
  mutate(csum = cumsum(ifelse(is.na(signed_angle_corrected), 0, signed_angle_corrected)) %>% round(4)) %>%
  ggplot(
    mapping = aes(x = rec_time, y = csum %>% scale())
  ) +
  geom_line(aes(col = stim_categ, group = 1)) +
  facet_wrap(~session, scales = "free_x", ncol = 1) +
  geom_label(data = text_layer, aes(x = -Inf, y = Inf, hjust = -.2, vjust = 1.3, label = paste0("Intensity: ", int_label))) +
  scale_color_manual(values = group_colors) +
  global_facet_theme
```

# Population analysis

## Rotation

### Calculating signed population angles

```{r message=FALSE, warning=FALSE, echo = FALSE}
corrected_angles_population <- coords_halo %>%
  dplyr::filter(
    body_part != "tail"
  ) %>%
  dplyr::group_by(animal_grp, session, frame) %>%
  mutate(
    vectx = diff(X_interp),
    vecty = diff(Y_interp),
  ) %>%
  group_by(animal_grp, session) %>%
  mutate(
    leadx = lead(vectx),
    leady = lead(vecty)
  ) %>%
  mutate(
    signed_angle = (atan2(leadx, leady) - atan2(vectx, vecty)) * 180 / (pi)
  ) %>%
  mutate(
    signed_angle_corrected = case_when(
      signed_angle >= 180 ~ 360 - signed_angle,
      signed_angle %>% abs() < 180 ~ signed_angle,
      signed_angle <= -180 ~ -360 - signed_angle,
      signed_angle > -180 | signed_angle < 0 ~ signed_angle
    )
  ) %>%
  ### removing unrealistically large angles
  mutate(
    signed_angle_corrected = ifelse(
      signed_angle_corrected %>% abs() > ANG_THRESHOLD,
      0,
      signed_angle_corrected
    )
  ) %>%
  ### adding ipsi or contra rotation variable
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::mutate(
    ipsicontra = case_when(
      sum(signed_angle_corrected, na.rm = T) > 0 & stim_side == "left" & stim_categ == "stim" ~ "ipsi",
      sum(signed_angle_corrected, na.rm = T) > 0 & stim_side == "right" & stim_categ == "stim" ~ "contra",
      sum(signed_angle_corrected, na.rm = T) < 0 & stim_side == "left" & stim_categ == "stim" ~ "contra",
      sum(signed_angle_corrected, na.rm = T) < 0 & stim_side == "right" & stim_categ == "stim" ~ "ipsi",
    )
  )
```

### Figure S7/B

```{r}
corrected_angles_population %>%
  dplyr::filter(
    body_part == "head",
    stim_categ != "no stim",
  ) %>%
  # dplyr::filter(intensity %in% c(SELECTED_INTENSITIES, "15 mW")) %>%
  dplyr::filter(intensity == "10 mW") %>%
  
  dplyr::group_by(animal_grp, stim_side, stim_categ) %>%
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected, na.rm = T),
    mean_ang = mean(signed_angle_corrected, na.rm = T),
    med_ang = median(signed_angle_corrected, na.rm = T),
    sd_ang = sd(signed_angle_corrected, na.rm = T),
    sem_ang = sem(signed_angle_corrected)
  ) %>%
  dplyr::mutate(
    ipsicontra = case_when(
      sum_ang > 0 & stim_side == "left" & stim_categ == "stim" ~ "ipsi",
      sum_ang > 0 & stim_side == "right" & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & stim_side == "left" & stim_categ == "stim" ~ "contra",
      sum_ang < 0 & stim_side == "right" & stim_categ == "stim" ~ "ipsi",
    )
  ) %>%
  dplyr::group_by(animal_grp) %>%
  fill(ipsicontra, .direction = "up") %>%
  ungroup() %>%
  dplyr::mutate(
    sum_ang_ipsicontra = ifelse(ipsicontra == "contra", sum_ang %>% abs(), -(sum_ang %>% abs())),
    mean_ang_ipsicontra = ifelse(ipsicontra == "contra", mean_ang %>% abs(), -(mean_ang %>% abs()))
  ) %>% # write.csv(file = file.path("output_data","to_publish","Emivel","PnO_mean_No_rotation_per_second_all_intensities_ipsicontra.csv"))
  ggplot(
    mapping = aes(
      x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
      y = (mean_ang_ipsicontra * FPS) / 360 # sum_ang %>% abs() /360  # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
    )
  ) +
  geom_point(aes(fill = animal_grp), shape = 21, size = 4) +
  geom_line(aes(group = animal_grp)) +
  geom_hline(yintercept = 0, col = "gray") +
  annotate(geom = "text", x = 3, hjust = -0.15, y = -0.02, label = "ipsilateral", col = "gray") +
  annotate(geom = "text", x = 3, hjust = -0.15, y = 0.02, label = "contralateral", col = "gray") +
  global_facet_theme +
  expand_limits(y = c(-0.2, 0.2), x = c(0, 4)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  ggtitle("Stimulus") +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )
```

### Figure S7/C

Mean/sum rotation angle by intensity (stimulus and control)

```{r message=FALSE, fig.width=15, fig.height=5, echo = FALSE}
plt1_angle <- corrected_angles_population %>%
  dplyr::filter(
    body_part == "head",
    stim_categ != "no stim",
    #stim_side == "left"
  ) %>%
  dplyr::filter(intensity %in% c(SELECTED_INTENSITIES, "15 mW")) %>%
  dplyr::group_by(animal_grp, stim_side, intensity, stim_categ) %>%
  dplyr::summarise(
    sum_ang = sum(signed_angle_corrected, na.rm = T),
    mean_ang = mean(signed_angle_corrected, na.rm = T),
    med_ang = median(signed_angle_corrected, na.rm = T),
    sd_ang = sd(signed_angle_corrected, na.rm = T),
    sem_ang = sem(signed_angle_corrected),
    all_stim_categ_length_row = length(signed_angle_corrected),
    mean_ang_manual = sum_ang / all_stim_categ_length_row
  ) %>% #write.csv(file = file.path("output_data","to_publish","Emivel","Halo_rotation.csv"))
  #dplyr::group_by(animal_grp, intensity) %>%
  ggplot(
    mapping = aes(
      x = forcats::fct_relevel(stim_categ, "before stim", "stim", "after stim"),
      # * 25: number of frames in a second. To get the mean rotation angle per second. /360: mean portion of one full circle per second
      y = ((mean_ang * FPS) %>% abs()) / 360 # %>% abs() /360  # sum_ang %>% abs() # / max(sum_ang %>% abs()) # %>% scale()
    )
  ) +
  geom_point(aes(fill = animal_grp), shape = 21, size = 4) +
  geom_line(aes(group = animal_grp)) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "less"),
    comparisons = list(c("before stim", "stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  ggpubr::stat_compare_means(
    method = "wilcox.test",
    paired = T,
    method.args = list(alternative = "greater"),
    comparisons = list(c("stim", "after stim")),
    vjust = 1.5,
    hide.ns = F
    # label.y = 3000
  ) +
  facet_grid(~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
  global_facet_theme +
  expand_limits(y = c(0, 4 * FPS / 360)) +
  # xlab("State") +
  # ylab("Scaled sum angle") +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black")
  )

plt1_angle +
  ggtitle(paste0(DOC_TITLE, ", ", "mean rotation angle by intensity"))

```




```{r}
generate_angle_dose_plot <- function(stimside) {
  corrected_angles_population %>%
    dplyr::filter(
      body_part == "head",
      stim_categ != "no stim",
      if (stimside == "right") {
        stim_side == stimside
      } else {
        stim_side == "left"
      }
    ) %>%
    dplyr::filter(intensity %in% c(SELECTED_INTENSITIES, "15 mW")) %>%
    dplyr::group_by(animal, intensity, stim_categ) %>%
    dplyr::summarise(
      sum_ang = sum(signed_angle_corrected, na.rm = T),
      mean_ang = mean(signed_angle_corrected, na.rm = T),
      med_ang = median(signed_angle_corrected, na.rm = T),
      sd_ang = sd(signed_angle_corrected, na.rm = T),
      sem_ang = sem(signed_angle_corrected)
    ) %>%
    dplyr::filter(
      stim_categ == "stim"
    ) %>% # write.csv(file = file.path("output_data","to_publish","Emivel","PnO_dose_intensity_rotation_control.csv"))

    ggplot(
      mapping = aes(
        x = forcats::fct_relevel(intensity, "1 mW", "5 mW", "10 mW", "15 mW"),
        y = ((mean_ang * FPS) %>% abs()) / 360
      )
    ) +
    geom_point(aes(fill = animal), shape = 21, size = 4) +
    geom_boxplot(alpha = 0) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = F,
      method.args = list(alternative = "two.sided"),
      comparisons = list(c("1 mW", "5 mW"), c("1 mW", "10 mW"), c("1 mW", "15 mW")),
      vjust = 1.5,
      hide.ns = F
      # label.y = 3000
    ) +
    global_facet_theme +
    expand_limits(y = c(0, 0.2)) +
    # xlab("State") +
    # ylab("Scaled sum angle") +
    # ggtitle("Stimulus") +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    )
}

# Generate plots for both control and non-control groups
plt_ang_dose_right <- generate_angle_dose_plot(stimside = "right")
plt_ang_dose_left <- generate_angle_dose_plot(stimside = "left")


cowplot::plot_grid(
  plt_ang_dose_right, plt_ang_dose_left, 
  ncol = 2,
  labels = paste0(DOC_TITLE, ", ", "mean rotation angle dose dependency")
)
```

```{r}
  corrected_angles_population %>%
    dplyr::filter(
      body_part == "head",
      stim_categ != "no stim"
    ) %>%
    dplyr::filter(intensity %in% c(SELECTED_INTENSITIES, "15 mW")) %>%
    dplyr::group_by(animal_grp, intensity, stim_categ) %>%
    dplyr::summarise(
      sum_ang = sum(signed_angle_corrected, na.rm = T),
      mean_ang = mean(signed_angle_corrected, na.rm = T),
      med_ang = median(signed_angle_corrected, na.rm = T),
      sd_ang = sd(signed_angle_corrected, na.rm = T),
      sem_ang = sem(signed_angle_corrected)
    ) %>%
    dplyr::filter(
      stim_categ == "stim"
    ) %>% # write.csv(file = file.path("output_data","to_publish","Emivel","Halo_dose_intensity_rotation.csv"))

    ggplot(
      mapping = aes(
        x = forcats::fct_relevel(intensity, "1 mW", "5 mW", "10 mW", "15 mW"),
        y = ((mean_ang * FPS) %>% abs()) / 360
      )
    ) +
    geom_point(aes(fill = animal_grp), shape = 21, size = 4) +
    geom_boxplot(alpha = 0) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = F,
      method.args = list(alternative = "two.sided"),
      comparisons = list(c("1 mW", "5 mW"), c("1 mW", "10 mW"), c("1 mW", "15 mW")),
      vjust = 1.5,
      hide.ns = F
      # label.y = 3000
    ) +
    global_facet_theme +
    expand_limits(y = c(0, 0.2)) +
    # xlab("State") +
    # ylab("Scaled sum angle") +
    # ggtitle("Stimulus") +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) +
  ggtitle(paste0(DOC_TITLE, ", ", "mean rotation angle dose dependency"))

```





## Traveled distance

```{r}
# saving the mean of the "before stim" "sum traveled distance" of all the animals (by intensity) for normalization

mean_sum_d_1mW <- coords_halo_1s_window %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0,
    body_part == "center",
    intensity %in% SELECTED_INTENSITIES
  ) %>%
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "1 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_5mW <- coords_halo_1s_window %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0,
    body_part == "center",
    intensity %in% SELECTED_INTENSITIES
  ) %>%
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "5 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_10mW <- coords_halo_1s_window %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0,
    body_part == "center",
    intensity %in% SELECTED_INTENSITIES
  ) %>%
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "10 mW"
  ) %>%
  pull(mean_sum_d)

mean_sum_d_15mW <- coords_halo_1s_window %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0,
    body_part == "center",
    intensity %in% c(SELECTED_INTENSITIES, "15 mW")
  ) %>%
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(mean_sum_d = mean(sum_d)) %>%
  dplyr::filter(
    stim_categ == "before stim",
    intensity == "15 mW"
  ) %>%
  pull(mean_sum_d)


coords_halo_1s_window %>%
  dplyr::filter(
    frame %% DATAPOINT_TO_PLOT == 0,
    body_part == "center",
    intensity %in% c(SELECTED_INTENSITIES, "15 mW")
  ) %>%
  dplyr::group_by(animal_grp, intensity, stim_categ) %>%
  dplyr::summarise(sum_d = sum(d, na.rm = T)) %>%
  dplyr::group_by(intensity, stim_categ) %>%
  dplyr::summarise(
    mean_sum_d = mean(sum_d),
    sd_sum_d = sd(sum_d)
  )
```



Combining 1 sec before stim and stim division to 1st and last 9 sec tables. In control cases if I normalize with the period length the "last_part" drops. The reason is that the animal moves at a constant rate so dividing the mean traveled distance will decrease the mean.

```{r}
generate_pop_dist_summary_plot <- function(stimside) {
  bind_rows(
    ### before stim 1 sec
    coords_halo_1s_window %>%
      dplyr::filter(
        stim_categ == "before stim",
        body_part == "center"
      ) %>%
      dplyr::mutate(
        stim_division = "before_stim",
        stim_div_length = 1
      ),
    ### stim devided to first_part (1s) and last part (9s)
    coords_halo %>%
      dplyr::filter(
        frame %% DATAPOINT_TO_PLOT == 0,
        stim_categ == "stim",
        body_part == "center"
      ) %>%
      ### recalculates distances after frames were removed
      dplyr::group_by(animal_grp, body_part, session) %>%
      dplyr::mutate(
        X_diff = c(diff(X_interp), NA) %>% round(4),
        Y_diff = c(diff(Y_interp), NA) %>% round(4),
        X_diff = ifelse(
          test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
            X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
          yes = 0,
          no = X_diff
        ),
        Y_diff = ifelse(
          Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
            Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
          yes = 0,
          no = Y_diff
        ),
        d = sqrt(X_diff^2 + Y_diff^2),
        d_mm = d / SCALE,
        speed_instant = d_mm / TIME_RES,
        mmps = speed_instant * (1 / TIME_RES),
        mps = (speed_instant * (1 / TIME_RES)) / 1000
      ) %>%
      dplyr::ungroup() %>%
      dplyr::group_by(animal_grp, body_part, session, stim_number) %>%
      dplyr::mutate(
        stim_division = ifelse(rec_time <= min(rec_time) + 1, yes = "first_part", no = "last_part"),
      )
  ) %>% 
    arrange(animal_grp, session, stim_number, rec_time) %>%
    dplyr::group_by(animal_grp, intensity, stim_categ, stim_division) %>%
    dplyr::mutate(
      stim_div_length_row = length(stim_division)
    ) %>%
    dplyr::group_by(
      animal_grp, intensity, stim_categ, stim_side, stim_division, across(stim_div_length_row)
    ) %>% 
    dplyr::summarise(
      sum_d = sum(d, na.rm = T),
      mean_d = mean(d, na.rm = T),
      max_sum_d = sum_d %>% max()
    ) %>% 
    dplyr::mutate(
      max_sum_d = replace(max_sum_d, stim_categ == "stim", NA),
      mean_sum_d = case_when(
        intensity == "1 mW" ~ mean_sum_d_1mW,
        intensity == "5 mW" ~ mean_sum_d_5mW,
        intensity == "10 mW" ~ mean_sum_d_10mW,
        intensity == "15 mW" ~ mean_sum_d_15mW,
      )
    ) %>% 
    dplyr::filter(
      if (stimside == "right") {
        stim_side == "right"
      } else {
        stim_side == "left"
      },
      intensity %in% c(SELECTED_INTENSITIES, "15 mW")
    ) %>% 
    ungroup() %>% 
    fill(max_sum_d, .direction = "down") %>% 
    ggplot(
      mapping = aes(
        x = forcats::fct_relevel(stim_division, "before_stim", "first_part", "last_part"),
        ### *6 to convert to mm/sec (6 rows represent one second)
        y = mean_d * 6 / SCALE # (sum_d/stim_div_length_row)/SCALE/10 # in [cm/sec] # mean_d/SCALE # / stim_div_length #/mean_sum_d
      )
    ) +
    geom_point(aes(fill = animal_grp %>% as.factor()), shape = 21, size = 4) +
    facet_wrap(~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique())), nrow = 1) +
    # facet_grid(animal_grp ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
    geom_line(aes(group = animal_grp)) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = T,
      comparisons = list(c("before_stim", "first_part")),
      method.args = list(alternative = "less"),
      vjust = 1.6,
      hide.ns = F
    ) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = T,
      comparisons = list(c("first_part", "last_part")),
      method.args = list(alternative = "greater"),
      vjust = 1.6,
      hide.ns = F
    ) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      label.y = 35 * 6,
      paired = T,
      comparisons = list(c("before_stim", "last_part")),
      method.args = list(alternative = "less"),
      vjust = 1.7,
      hide.ns = F
    ) +
    expand_limits(y = c(0, 40 * 6)) +
    global_facet_theme +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) +
    if (stimside == "right") {
      ggtitle("Right side")
    } else {
      ggtitle("Left side")
    }
}

plt_pop_dist_sum_left <- generate_pop_dist_summary_plot(stimside = "left")
plt_pop_dist_sum_right <- generate_pop_dist_summary_plot(stimside = "right")

cowplot::plot_grid(
  plt_pop_dist_sum_right, plt_pop_dist_sum_left, 
  nrow = 2,
  labels = paste0(DOC_TITLE, ", " , "traveled distance")
  )
```

### Figure S7/E

```{r fig.width=15}
  bind_rows(
    ### before stim 1 sec
    coords_halo_1s_window %>%
      dplyr::filter(
        stim_categ == "before stim",
        body_part == "center"
      ) %>%
      dplyr::mutate(
        stim_division = "before_stim",
        stim_div_length = 1
      ),
    ### stim devided to first_part (1s) and last part (9s)
    coords_halo %>%
      dplyr::filter(
        frame %% DATAPOINT_TO_PLOT == 0,
        stim_categ == "stim",
        body_part == "center"
      ) %>%
      ### recalculates distances after frames were removed
      dplyr::group_by(animal_grp, body_part, session) %>%
      dplyr::mutate(
        X_diff = c(diff(X_interp), NA) %>% round(4),
        Y_diff = c(diff(Y_interp), NA) %>% round(4),
        X_diff = ifelse(
          test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
            X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
          yes = 0,
          no = X_diff
        ),
        Y_diff = ifelse(
          Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
            Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
          yes = 0,
          no = Y_diff
        ),
        d = sqrt(X_diff^2 + Y_diff^2),
        d_mm = d / SCALE,
        speed_instant = d_mm / TIME_RES,
        mmps = speed_instant * (1 / TIME_RES),
        mps = (speed_instant * (1 / TIME_RES)) / 1000
      ) %>%
      dplyr::ungroup() %>%
      dplyr::group_by(animal_grp, body_part, session, stim_number) %>%
      dplyr::mutate(
        stim_division = ifelse(rec_time <= min(rec_time) + 1, yes = "first_part", no = "last_part"),
      )
  ) %>%
    arrange(animal_grp, session, stim_number, rec_time) %>%
    dplyr::group_by(animal_grp, intensity, stim_categ, stim_division) %>%
    dplyr::mutate(
      stim_div_length_row = length(stim_division)
    ) %>%
    dplyr::group_by(
      animal_grp, intensity, stim_categ, stim_side, stim_division, across(stim_div_length_row)
    ) %>%
    dplyr::summarise(
      sum_d = sum(d, na.rm = T),
      mean_d = mean(d, na.rm = T),
      max_sum_d = sum_d %>% max()
    ) %>%
    dplyr::mutate(
      max_sum_d = replace(max_sum_d, stim_categ == "stim", NA),
      mean_sum_d = case_when(
        intensity == "1 mW" ~ mean_sum_d_1mW,
        intensity == "5 mW" ~ mean_sum_d_5mW,
        intensity == "10 mW" ~ mean_sum_d_10mW,
        intensity == "15 mW" ~ mean_sum_d_15mW,
      )
    ) %>%
    dplyr::filter(
      intensity %in% c(SELECTED_INTENSITIES, "15 mW")
    ) %>%
    ungroup() %>%
    fill(max_sum_d, .direction = "down") %>% #write.csv(file = file.path("output_data","to_publish","Emivel","Halo_traveled_distance.csv"))
    ggplot(
      mapping = aes(
        x = forcats::fct_relevel(stim_division, "before_stim", "first_part", "last_part"),
        ### *6 to convert to mm/sec (6 rows represent one second)
        y = mean_d * 6 / SCALE # (sum_d/stim_div_length_row)/SCALE/10 # in [cm/sec] # mean_d/SCALE # / stim_div_length #/mean_sum_d
      )
    ) +
    geom_point(aes(fill = animal_grp %>% as.factor()), shape = 21, size = 4) +
    facet_wrap(~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique())), nrow = 1) +
    # facet_grid(animal_grp ~ factor(intensity, levels = gtools::mixedsort(intensity %>% unique()))) +
    geom_line(aes(group = animal_grp)) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = T,
      comparisons = list(c("before_stim", "first_part")),
      method.args = list(alternative = "less"),
      vjust = 1.6,
      hide.ns = F
    ) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = T,
      comparisons = list(c("first_part", "last_part")),
      method.args = list(alternative = "greater"),
      vjust = 1.6,
      hide.ns = F
    ) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      label.y = 35 * 6,
      paired = T,
      comparisons = list(c("before_stim", "last_part")),
      method.args = list(alternative = "less"),
      vjust = 1.7,
      hide.ns = F
    ) +
    expand_limits(y = c(0, 40 * 6)) +
    global_facet_theme +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) + 
    ggtitle(paste0(DOC_TITLE, ", " , "traveled distance"))
```



```{r}
generate_distance_dose_plot <- function(stimside) {
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "stim",
      body_part == "center"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
   
    dplyr::filter(
      if (stimside == "right") {
        stim_side == stimside
      } else {
        stim_side == "left"
      },
      intensity %in% c(SELECTED_INTENSITIES, "15 mW")
    ) %>%
    dplyr::group_by(animal, intensity) %>% # dplyr::filter(animal == "GIIFM18b", intensity == "1 mW") %>% pull(d) %>% sum(na.rm = T) / 250
    # dplyr::mutate(nrows=length(intensity)) %>%
    dplyr::summarise(
      mean_d = mean(d, na.rm = T),
      sum_d_norm = sum(d, na.rm = T) / 250,
      nrows = length(intensity)
    ) %>% # write.csv(file = file.path("output_data","to_publish","Emivel","dose_intensity_diff_trav_dist.csv"))
    ggplot(
      aes(x = forcats::fct_relevel(intensity, "1 mW", "5 mW", "10 mW", "15 mW"), y = mean_d * 6)
    ) +
    geom_point(aes(fill = animal), shape = 21, size = 4) +
    geom_boxplot(alpha = 0) +
    expand_limits(y = c(0, 60 * 6)) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = F,
      comparisons = list(c("1 mW", "5 mW"), c("1 mW", "10 mW"), c("1 mW", "15 mW")),
      method.args = list(alternative = "two.sided"),
      vjust = 1.6,
      hide.ns = F
    ) +
    global_facet_theme +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) +
    if (stimside == "right") {
      ggtitle("Right side")
    } else {
      ggtitle("Left side")
    }
}

plt_distance_dose_right <- generate_distance_dose_plot(stimside = "right")
plt_distance_dose_left <- generate_distance_dose_plot(stimside = "left")


cowplot::plot_grid(
  plt_distance_dose_right, plt_distance_dose_left, 
  ncol = 2,
  labels = paste0(DOC_TITLE, ", " , "traveled distance dose dependency")
  )
```

```{r}
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "stim",
      body_part == "center"
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
   
    dplyr::filter(
      intensity %in% c(SELECTED_INTENSITIES, "15 mW")
    ) %>%
    dplyr::group_by(animal_grp, intensity) %>% # dplyr::filter(animal == "GIIFM18b", intensity == "1 mW") %>% pull(d) %>% sum(na.rm = T) / 250
    # dplyr::mutate(nrows=length(intensity)) %>%
    dplyr::summarise(
      mean_d = mean(d, na.rm = T),
      sum_d_norm = sum(d, na.rm = T) / 250,
      nrows = length(intensity)
    ) %>% #write.csv(file = file.path("output_data","to_publish","Emivel","Halo_dose_intensity_trav_dist.csv"))
    ggplot(
      aes(x = forcats::fct_relevel(intensity, "1 mW", "5 mW", "10 mW", "15 mW"), y = mean_d * 6)
    ) +
    geom_point(aes(fill = animal_grp), shape = 21, size = 4) +
    geom_boxplot(alpha = 0) +
    expand_limits(y = c(0, 60 * 6)) +
    ggpubr::stat_compare_means(
      method = "wilcox.test",
      paired = F,
      comparisons = list(c("1 mW", "5 mW"), c("1 mW", "10 mW"), c("1 mW", "15 mW")),
      method.args = list(alternative = "two.sided"),
      vjust = 1.6,
      hide.ns = F
    ) +
    global_facet_theme +
    theme(aspect.ratio = 1) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.text = element_text(size = 12),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black")
    ) +
  ggtitle(paste0(DOC_TITLE, ", " , "traveled distance dose dependency"))
```

```{r}
bind_rows(
  ### before stim 1 sec
  coords_halo_1s_window %>%
    dplyr::filter(
      stim_categ == "before stim",
      body_part == "center"
    ) %>%
    dplyr::mutate(
      stim_division = "before_stim",
      stim_div_length = 1
    ),
  ### stim devided to first_part (1s) and last part (9s)
  coords_halo %>%
    dplyr::filter(
      frame %% DATAPOINT_TO_PLOT == 0,
      stim_categ == "stim",
      body_part == "center"
      # animal %!in% CONTROL_ANIMALS,
      # body_part == "center",
      # intensity %in% SELECTED_INTENSITIES
    ) %>%
    ### recalculates distances after frames were removed
    dplyr::group_by(animal_grp, body_part, session) %>%
    dplyr::mutate(
      X_diff = c(diff(X_interp), NA) %>% round(4),
      Y_diff = c(diff(Y_interp), NA) %>% round(4),
      X_diff = ifelse(
        test = X_diff < quantile(X_diff, na.rm = T)["25%"] - 3 * IQR(X_diff, na.rm = T) |
          X_diff > quantile(X_diff, na.rm = T)["75%"] + 3 * IQR(X_diff, na.rm = T),
        yes = 0,
        no = X_diff
      ),
      Y_diff = ifelse(
        Y_diff < quantile(Y_diff, na.rm = T)["25%"] - 3 * IQR(Y_diff, na.rm = T) |
          Y_diff > quantile(Y_diff, na.rm = T)["75%"] + 3 * IQR(Y_diff, na.rm = T),
        yes = 0,
        no = Y_diff
      ),
      d = sqrt(X_diff^2 + Y_diff^2),
      d_mm = d / SCALE,
      speed_instant = d_mm / TIME_RES,
      mmps = speed_instant * (1 / TIME_RES),
      mps = (speed_instant * (1 / TIME_RES)) / 1000
    ) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(animal_grp, body_part, session, stim_number) %>%
    dplyr::mutate(
      stim_division = ifelse(rec_time <= min(rec_time) + 1, yes = "first_part", no = "last_part"),
      stim_div_length = ifelse(rec_time <= min(rec_time) + 1, yes = 1, no = 9)
    )
) %>%
  dplyr::group_by(animal_grp, body_part, session, stim_number, stim_division) %>%
  # dplyr::mutate(stim_div_length = length(stim_division)) %>%
  ungroup() %>%
  arrange(animal_grp, session, stim_number, rec_time) %>%
  dplyr::group_by(animal_grp, body_part, intensity, stim_categ) %>% # stim_number,
  dplyr::summarise(
    sum_d = sum(d, na.rm = T),
    mean_d = mean(d, na.rm = T),
    max_sum_d = sum_d %>% max()
  ) %>%
  dplyr::mutate(
    max_sum_d = replace(max_sum_d, stim_categ == "stim", NA),
    mean_sum_d = case_when(
      intensity == "1 mW" ~ mean_sum_d_1mW,
      intensity == "5 mW" ~ mean_sum_d_5mW,
      intensity == "10 mW" ~ mean_sum_d_10mW,
      intensity == "15 mW" ~ mean_sum_d_15mW,
    )
  ) %>%
  dplyr::filter(
    intensity %in% c(SELECTED_INTENSITIES, "15 mW"),
    # manually removing one line causing problem (in GIIFM20b 15 mW stim was repeated in two sessions, session 6 was not following protocol)
  ) %>%
  ungroup() %>%
  fill(max_sum_d, .direction = "down") %>%
  dplyr::filter(
    intensity %in% c(SELECTED_INTENSITIES, "15 mW"),
    stim_categ == "stim"
  )
# %>% View()
#   dplyr::group_by(animal, intensity) %>%
#   dplyr::summarise(
#     mean_d = mean(d,na.rm = T)
# ) %>%
```

